
---
output:
  html_document: default
  pdf_document: default
---

# Reproducible analysis of the faculty AI for data science survey


```{r setup-faculty-analysis, warning=FALSE, message=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
  dplyr, stringr, ggplot2, knitr, scales, janitor, forcats, tidyr,
  readxl, tibble, purrr
)

# Create output directory for saved figures
if (!dir.exists("figures_all")) dir.create("figures_all", recursive = TRUE)

# Helper to locate the faculty survey workbook inside the repository
find_faculty_file <- function() {
  candidates <- c("AI4D_FacultySurvey_August_8_2025.xlsx",
                  file.path("data","AI4D_FacultySurvey_August_8_2025.xlsx")
  )
  found <- candidates[file.exists(candidates)]
  if (length(found) == 0) {
    stop(
      "Faculty survey Excel file not found in the working directory. ",
      "Place the workbook in the repo root or update `find_faculty_file()`."
    )
  }
  found[[1]]
}
```

```{r import-faculty-data, warning=FALSE, message=FALSE}
# Read the workbook using repo-relative paths only
faculty_data_path <- find_faculty_file()

# ---- 1) Read the first two rows (Qualtrics short & long headers) ----
hdr <- readxl::read_excel(faculty_data_path, n_max = 2, col_names = FALSE)

r1 <- as.character(unlist(hdr[1, ], use.names = FALSE))
r2 <- as.character(unlist(hdr[2, ], use.names = FALSE))

# Heuristic to pick which row is the short (system) header
score_row <- function(x) {
  keys <- c("StartDate","EndDate","ResponseId","Progress","Duration (in seconds)")
  sum(keys %in% x) + sum(stringr::str_detect(x, "^Q\\d"))
}
short_raw <- if (score_row(r1) >= score_row(r2)) r1 else r2
long_raw  <- if (identical(short_raw, r1)) r2 else r1

# Ensure consistent length, clean short names for valid R columns
n_cols <- max(length(short_raw), length(long_raw))
short_clean <- janitor::make_clean_names(short_raw[seq_len(n_cols)])
long_clean  <- long_raw[seq_len(n_cols)]

# Map short -> long labels for nice printing later
label_map <- setNames(long_clean, short_clean)

# ---- 2) Read full data using the short names; data begin on row 3 ----
faculty_data <- readxl::read_excel(
  faculty_data_path,
  sheet = 1,
  skip = 2,
  col_names = short_clean,
  guess_max = 10000
) |>
  janitor::remove_empty("rows") |>
  janitor::remove_empty("cols")

attr(faculty_data, "label_map") <- label_map

# ---- 3) Optional quality filters ----
faculty_data <- faculty_data |>
  dplyr::mutate(
    duration_in_seconds = suppressWarnings(as.numeric(duration_in_seconds)),
    progress            = suppressWarnings(as.numeric(progress))
  )

if (all(c("duration_in_seconds","progress") %in% names(faculty_data))) {
  faculty_data <- faculty_data |>
    dplyr::filter(duration_in_seconds >= 180, progress >= 25)
}

cat("Rows after quality filters:", nrow(faculty_data), "
Columns:", ncol(faculty_data), "
")
dplyr::glimpse(faculty_data)

survey_palette_red_teal <- c("#117A65", "#48C9B0", "#F1C40F", "#E67E22", "#8B0000")
```


## Q2: Consent to Participate in the Study

```{r, warning=FALSE, message=FALSE}

# ---- Locate Q2 column (handles 'Q2' or 'q2') ----
q2_col <- dplyr::case_when(
  "Q2" %in% names(faculty_data) ~ "Q2",
  "q2" %in% names(faculty_data) ~ "q2",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q2_col))

# ---- Standardize Q2 to two levels; handle text or numeric exports ----
faculty_data <- faculty_data %>%
  mutate(
    q2_raw = .data[[q2_col]],
    q2_std = case_when(
      is.numeric(q2_raw) & q2_raw == 1 ~ "Consented",
      is.numeric(q2_raw) & q2_raw == 2 ~ "Did Not Consent",
      str_detect(q2_raw, regex("^i\\s*voluntar|consent to participate",
                               ignore_case = TRUE)) ~ "Consented",
      str_detect(q2_raw, regex("don.?t\\s*consent|do not consent|decline",
                               ignore_case = TRUE)) ~ "Did Not Consent",
      TRUE ~ NA_character_
    ) |> factor(levels = c("Consented", "Did Not Consent"))
  )

# ---- Summary table ----
q2_tab <- faculty_data %>%
  filter(!is.na(q2_std)) %>%
  count(q2_std, name = "Count") %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

knitr::kable(q2_tab, caption = "Q2: Consent Summary")

# ---- Plots (keep SAME color as student survey: steelblue) ----
ggplot(q2_tab, aes(x = q2_std, y = Percent)) +
  geom_col(fill = "#48C9B0") +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(title = "Q2: Consent (Percent)", x = NULL, y = "Percent") +
  theme_minimal()

# ---- Create consented-only dataset for downstream questions ----
faculty_data_consented <- faculty_data %>% filter(q2_std == "Consented")
cat("Consented N:", nrow(faculty_data_consented), "of", nrow(faculty_data), "total\n")


```


## Saving custom theme for future use
```{r}
# --- Define once at the top of your script or in a separate R file ---
theme_publication <- function(base_size = 10) {
  theme_classic(base_size = base_size) +
    theme(
      plot.title      = element_text(face = "bold", size = 11, hjust = 0.5),
      plot.subtitle   = element_text(size = 10, hjust = 0.5, color = "gray25"),
      axis.text.x     = element_text(face = "bold", size = 10),
      axis.text.y     = element_text(face = "bold", size = 10, lineheight = 1.1),
      axis.title.x    = element_text(face = "bold", size = 10),
      axis.title.y    = element_text(face = "bold", size = 10),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor = element_blank(),
      legend.title    = element_text(face = "bold"),
      legend.text     = element_text(size = 8),
      legend.position = "bottom",
      legend.key.size = grid::unit(0.35, "cm"),
      legend.margin   = margin(t = 2, b = 5),
      legend.box.margin = margin(5, 5, 10, 5),
      legend.box = "vertical",
      plot.caption    = element_text(size = 9, hjust = 0, color = "gray30"),
      plot.margin     = margin(10, 15, 45, 15)
    )
}

###
load_publication_style <- function() {
  # Color palette
  assign("familiarity_palette", c(
    "Not at All Familiar" = "#B2182B",
    "Not Very Familiar"   = "#E67E22",
    "Neutral"             = "#F1C40F",
    "Somewhat Familiar"   = "#48C9B0",
    "Very Familiar"       = "#117A65"
  ), envir = .GlobalEnv)

  # Theme
  theme_set(theme_publication())
}

load_publication_style()


```


## Q3: Gender-summary

```{r, warning=FALSE, message=FALSE}

# Use consented data if it exists; otherwise fall back to full
faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# Locate Q3 column robustly
q3_col <- dplyr::case_when(
  "Q3" %in% names(faculty_analysis_data) ~ "Q3",
  "q3" %in% names(faculty_analysis_data) ~ "q3",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q3_col))

# Standardize Q3 values (handles numeric-coded or text exports)
faculty_analysis_data <- faculty_analysis_data %>%
  mutate(
    q3_raw = .data[[q3_col]],
    q3_std = case_when(
      is.numeric(q3_raw) & q3_raw == 1 ~ "Male",
      is.numeric(q3_raw) & q3_raw == 2 ~ "Female",
      is.numeric(q3_raw) & q3_raw == 3 ~ "Prefer not to say",
      str_detect(q3_raw, regex("^male$", ignore_case = TRUE)) ~ "Male",
      str_detect(q3_raw, regex("^female$", ignore_case = TRUE)) ~ "Female",
      str_detect(q3_raw, regex("prefer\\s*not\\s*to\\s*say",
                               ignore_case = TRUE)) ~ "Prefer not to say",
      TRUE ~ NA_character_
    ) |> factor(levels = c("Male","Female","Prefer not to say"))
  )

# Summary table
q3_tab <- faculty_analysis_data %>%
  filter(!is.na(q3_std)) %>%
  count(q3_std, name = "Count") %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

knitr::kable(q3_tab, caption = "Q3: Gender — Summary")

# Plots
ggplot(q3_tab, aes(x = q3_std, y = Percent)) +
  geom_col(fill = "#48C9B0") +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(title = "Q3: Gender — Percent", x = NULL, y = "Percent") +
  theme_minimal()

```


## Q4: What is your primary area of expertise within data science? (e.g., Intro Statistics, Mathematical Statistics, Data Science, Machine Learning, etc.)

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q4_col <- dplyr::case_when(
  "Q4" %in% names(faculty_analysis_data) ~ "Q4",
  "q4" %in% names(faculty_analysis_data) ~ "q4",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q4_col))

# ---- Clean & map free text into canonical categories ----
# We keep the FIRST matching category as the "primary" area.
canon <- faculty_analysis_data %>%
  transmute(
    q4_raw  = as.character(.data[[q4_col]]),
    q4_norm = str_squish(str_to_lower(q4_raw)),
    area = case_when(
      str_detect(q4_norm, "intro.*stat|^intro\\b")                              ~ "Introductory Statistics",
      str_detect(q4_norm, "mathematical stat|theoretical stat|probability(?!.*appl)") ~ "Mathematical Statistics / Probability",
      str_detect(q4_norm, "\\bdata science\\b|\\bds\\b")                        ~ "Data Science",
      str_detect(q4_norm, "machine learning|\\bml\\b")                          ~ "Machine Learning",
      str_detect(q4_norm, "deep learning|\\bdl\\b")                              ~ "Deep Learning",
      str_detect(q4_norm, "biostat")                                            ~ "Biostatistics",
      str_detect(q4_norm, "applied stat")                                       ~ "Applied Statistics",
      str_detect(q4_norm, "econometric|business analytics|analytics\\b")        ~ "Business Analytics / Econometrics",
      str_detect(q4_norm, "visualiz|viz")                                       ~ "Data Visualization",
      str_detect(q4_norm, "computer science|\\bcs\\b|software")                 ~ "Computer Science / Software",
      str_detect(q4_norm, "programming|python|\\br\\b|java|matlab|c\\+\\+")     ~ "Programming / Tools",
      str_detect(q4_norm, "database|data engineer|etl|pipeline")                ~ "Data Engineering / Databases",
      str_detect(q4_norm, "big data|spark|hadoop|aws|azure|gcp|cloud")          ~ "Big Data / Cloud",
      str_detect(q4_norm, "nlp|natural language")                                ~ "Natural Language Processing",
      str_detect(q4_norm, "optimiza|operations research|\\bor\\b")              ~ "Optimization / Operations Research",
      str_detect(q4_norm, "statistic")                                          ~ "Statistics (General)",
      str_length(q4_norm) > 0                                                   ~ "Other / Unspecified",
      TRUE                                                                      ~ NA_character_
    )
  )

# ---- Summary table ----
q4_tab <- canon %>%
  filter(!is.na(area)) %>%
  count(area, name = "Count") %>%
  arrange(desc(Count)) %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

knitr::kable(q4_tab, caption = "Q4: Primary Area of Expertise — Summary")

# ---- Plots (steelblue; flip for readability) ----
# Wrap long labels
q4_tab <- q4_tab %>% mutate(area_lab = stringr::str_wrap(area, width = 28))

p_q4<-ggplot(q4_tab, aes(x = fct_reorder(area_lab, Count), y = Count)) +
  geom_col(fill = "#48C9B0") +
  coord_flip() +
  labs(title = "Q4: Primary Area of Expertise", x = NULL, y = "Count") +
  theme_minimal()

ggsave(
  filename = "Faculty_figures/Q4_Primary Area of Expertise.png",
  plot = p_q4,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)
```

## Q5: How many years have you been teaching data science or related subjects?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q5_col <- dplyr::case_when(
  "Q5" %in% names(faculty_analysis_data) ~ "Q5",
  "q5" %in% names(faculty_analysis_data) ~ "q5",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q5_col))

# Map numeric or text to canonical labels
q5_raw <- faculty_analysis_data[[q5_col]]
q5_num <- suppressWarnings(as.numeric(q5_raw))
levels_q5 <- c("1–3 years","3–5 years","5–10 years","10+ years")

faculty_analysis_data <- faculty_analysis_data %>%
  mutate(
    q5_std = case_when(
      !is.na(q5_num) & q5_num == 1 ~ "1–3 years",
      !is.na(q5_num) & q5_num == 2 ~ "3–5 years",
      !is.na(q5_num) & q5_num == 3 ~ "5–10 years",
      !is.na(q5_num) & q5_num == 4 ~ "10+ years",
      str_detect(as.character(q5_raw), regex("^1\\s*[-–]\\s*3",
                                             ignore_case = TRUE)) ~ "1–3 years",
      str_detect(as.character(q5_raw), regex("^3\\s*[-–]\\s*5",
                                             ignore_case = TRUE)) ~ "3–5 years",
      str_detect(as.character(q5_raw), regex("^5\\s*[-–]\\s*10",
                                             ignore_case = TRUE)) ~ "5–10 years",
      str_detect(as.character(q5_raw), regex("^10\\+|10\\s*\\+|10\\s*\\+\\s*years?",
                                             ignore_case = TRUE)) ~ "10+ years",
      TRUE ~ NA_character_
    ) |> factor(levels = levels_q5)
  )

# Summary table
q5_tab <- faculty_analysis_data %>%
  filter(!is.na(q5_std)) %>%
  count(q5_std, name = "Count") %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

knitr::kable(q5_tab, caption = "Q5: Years Teaching DS or Related Subjects — Summary")

# Bar charts (counts & percent)
ggplot(q5_tab, aes(x = q5_std, y = Count)) +
  geom_col(fill = "#48C9B0") +
  labs(title = "Q5: Years Teaching DS or Related Subjects", x = NULL, y = "Count") +
  theme_minimal()

```


## Q6: At what academic level do you primarily teach?

```{r, warning=FALSE, message=FALSE}
faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q6_col <- dplyr::case_when(
  "Q6" %in% names(faculty_analysis_data) ~ "Q6",
  "q6" %in% names(faculty_analysis_data) ~ "q6",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q6_col))

# Map numeric or text to canonical labels
q6_raw <- faculty_analysis_data[[q6_col]]
q6_num <- suppressWarnings(as.numeric(q6_raw))
levels_q6 <- c("Undergraduate","Graduate","Both")

faculty_analysis_data <- faculty_analysis_data %>%
  mutate(
    q6_std = case_when(
      !is.na(q6_num) & q6_num == 1 ~ "Undergraduate",
      !is.na(q6_num) & q6_num == 2 ~ "Graduate",
      !is.na(q6_num) & q6_num == 3 ~ "Both",
      str_detect(as.character(q6_raw), regex("^undergrad|^undergraduate",
                                             ignore_case = TRUE)) ~ "Undergraduate",
      str_detect(as.character(q6_raw), regex("^grad|^graduate$", ignore_case = TRUE)) &
        !str_detect(as.character(q6_raw), regex("undergrad|both", ignore_case = TRUE)) ~ "Graduate",
      str_detect(as.character(q6_raw), regex("\\bboth\\b|ug\\s*&?\\s*grad|undergrad.*grad|undergraduate.*graduate",
                                             ignore_case = TRUE)) ~ "Both",
      TRUE ~ NA_character_
    ) |> factor(levels = levels_q6)
  )

# Summary table
q6_tab <- faculty_analysis_data %>%
  filter(!is.na(q6_std)) %>%
  count(q6_std, name = "Count") %>%
  mutate(Percent = round(100 * Count / sum(Count), 1))

knitr::kable(q6_tab, caption = "Q6: Primary Teaching Level — Summary")

# Plots (steelblue to match student survey)
ggplot(q6_tab, aes(x = q6_std, y = Count)) +
  geom_col(fill = "#48C9B0") +
  labs(title = "Q6: Primary Teaching Level", x = NULL, y = "Count") +
  theme_minimal()


```

## Q7: Have you incorporated AI tools in your data science teaching?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q7_col <- dplyr::case_when(
  "Q7" %in% names(faculty_analysis_data) ~ "Q7",
  "q7" %in% names(faculty_analysis_data) ~ "q7",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q7_col))

# Map numeric or text to canonical labels
q7_raw <- faculty_analysis_data[[q7_col]]
q7_num <- suppressWarnings(as.numeric(q7_raw))
levels_q7 <- c("Never","Rarely","Sometimes","Often","Always")

faculty_analysis_data <- faculty_analysis_data %>%
  mutate(
    q7_std = case_when(
      !is.na(q7_num) & q7_num == 1 ~ "Never",
      !is.na(q7_num) & q7_num == 2 ~ "Rarely",
      !is.na(q7_num) & q7_num == 3 ~ "Sometimes",
      !is.na(q7_num) & q7_num == 4 ~ "Often",
      !is.na(q7_num) & q7_num == 5 ~ "Always",
      str_detect(as.character(q7_raw), regex("^never$", ignore_case = TRUE)) ~ "Never",
      str_detect(as.character(q7_raw), regex("^rarely$", ignore_case = TRUE)) ~ "Rarely",
      str_detect(as.character(q7_raw), regex("^sometimes?$", ignore_case = TRUE)) ~ "Sometimes",
      str_detect(as.character(q7_raw), regex("^often$", ignore_case = TRUE)) ~ "Often",
      str_detect(as.character(q7_raw), regex("^always$", ignore_case = TRUE)) ~ "Always",
      TRUE ~ NA_character_
    ) |> factor(levels = levels_q7)
  )


```


## Q7 counts and percents
```{r Q7}
# ---- 1) Summarize counts & percentages ----
q7_tab <- faculty_analysis_data %>%
  filter(!is.na(q7_std) & q7_std != "") %>%       # valid (non-empty) responses
  count(q7_std, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Label = paste0(Count, " (", round(Percent * 100, 1), "%)")
  )

n_valid <- sum(q7_tab$Count)  # total valid responses

# ---- 2) Horizontal bar chart (publication style) ----
p_q7 <- ggplot(q7_tab, aes(
  y = fct_reorder(q7_std, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels showing counts and percents
  geom_text(
    aes(label = Label),
    hjust = -0.1, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(q7_tab$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles & captions
  labs(
    title = "Q7: Incorporated AI Tools in Teaching",
    subtitle = "Percentages based on valid (non-missing) faculty responses",
    x = "Percentage of Faculty",
    y = "Response",
    caption = paste0(
      "n = ", n_valid,
      " faculty respondents. (Percentages calculated from valid responses.)"
    )
  ) +

  # Publication-style theme (matching Q22)
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin = margin(10, 25, 25, 15)
  )

# ---- 3) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q7_Used_AI_horizontal_pub.png",
  plot = p_q7,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

# ---- 4) Display plot and summary table ----
print(p_q7)
knitr::kable(q7_tab, caption = paste0("Q7: Incorporated AI Tools in Teaching (n = ", n_valid, ")"))
```


## Q8: What AI tools or platforms do you use in your teaching?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# ---- Canonical labels in the instrument order ----
opt_map <- c(
  "1"="AI coding platforms \n(Jupyter/Colab)",
  "2"="Automated grading",
  "3"="AI learning assistants \n(ChatGPT)",
  "4"="AI data analysis/ vizualizations \n(DataRobot/Tableau)",
  "5"="AI project management/ \ncollaboration",
  "6"="Other",
  "7"="None"
)
canon_levels <- unname(opt_map)

# ---- Detect export style ----
nm <- names(faculty_analysis_data)
q8_single <- nm[tolower(nm) == "q8"]
q8_wide   <- nm[str_detect(tolower(nm), "^q8(\\.|_|\\s|\\[|:)|^q8\\d")]
q8_wide   <- setdiff(q8_wide, nm[str_detect(tolower(nm), "text|comment")])

faculty_analysis_data$row_id <- seq_len(nrow(faculty_analysis_data))

if (length(q8_single) == 1 && length(q8_wide) == 0) {
  # ---- Case A: single comma/semicolon-separated column ----
  long <- faculty_analysis_data %>%
    select(row_id, q8 = all_of(q8_single)) %>%
    filter(!is.na(q8), q8 != "") %>%
    mutate(q8 = str_split(q8, "[,;]")) %>%
    unnest(q8) %>%
    mutate(q8 = str_squish(q8),
           option = case_when(
             str_detect(q8, regex("coding|jupyter|colab",  TRUE)) ~ opt_map["1"],
             str_detect(q8, regex("grading",                TRUE)) ~ opt_map["2"],
             str_detect(q8, regex("assistant|chatgpt",      TRUE)) ~ opt_map["3"],
             str_detect(q8, regex("analysis|visual|datarobot|tableau", TRUE)) ~ opt_map["4"],
             str_detect(q8, regex("project|collab",         TRUE)) ~ opt_map["5"],
             str_detect(q8, regex("^none$",                 TRUE)) ~ opt_map["7"],
             str_detect(q8, regex("other|specify",          TRUE)) ~ opt_map["6"],
             TRUE ~ NA_character_
           )) %>%
    filter(!is.na(option))
} else if (length(q8_wide) >= 1) {
  # ---- Case B: multiple columns Q8_1..Q8_7 (or labeled variants) ----
  long <- faculty_analysis_data %>%
    select(row_id, all_of(q8_wide)) %>%
    pivot_longer(-row_id, names_to = "col", values_to = "val") %>%
    filter(!(is.na(val) | val %in% c("", "0", "Not Selected", "FALSE", "False"))) %>%
    mutate(
      idx = str_extract(col, "\\d+"),
      option = unname(opt_map[idx]),
      option = coalesce(option, case_when(
        str_detect(col, regex("coding|jupyter|colab",  TRUE)) ~ opt_map["1"],
        str_detect(col, regex("grading",                TRUE)) ~ opt_map["2"],
        str_detect(col, regex("assistant|chatgpt",      TRUE)) ~ opt_map["3"],
        str_detect(col, regex("analysis|visual|datarobot|tableau", TRUE)) ~ opt_map["4"],
        str_detect(col, regex("project|collab",         TRUE)) ~ opt_map["5"],
        str_detect(col, regex("\\bnone\\b",             TRUE)) ~ opt_map["7"],
        str_detect(col, regex("other|specify",          TRUE)) ~ opt_map["6"],
        TRUE ~ NA_character_
      ))
    ) %>%
    filter(!is.na(option))
} else {
  stop("Q8 columns not found in this export.")
}

# ---- Enforce 'None' exclusivity ----
long <- long %>%
  group_by(row_id) %>% mutate(has_other = any(option != "None")) %>%
  ungroup() %>% filter(!(option == "None" & has_other)) %>%
  select(-has_other)


```


## Q8 plot
```{r}
# ---- 1) Summarize counts and percentages ----

n_valid <- 14  # total number of valid (non-empty) responses
q8_tab <- long %>%
  count(option, name = "Count") %>%
  complete(option = canon_levels, fill = list(Count = 0)) %>%
  mutate(
    option = factor(option, levels = canon_levels),
    Percent = Count / n_valid,
    Label = paste0(Count, " (", round(Percent * 100, 1), "%)")
  )

# ---- 2) Wrap long labels for readability ----
opt_wrapped <- stringr::str_wrap(canon_levels, width = 32)
plot_data <- q8_tab %>%
  mutate(option_w = factor(stringr::str_wrap(as.character(option), 32),
                           levels = rev(opt_wrapped)))

# ---- 3) Horizontal bar chart (publication style) ----
p_q8F <- ggplot(plot_data, aes(
  y = fct_reorder(option_w, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels showing counts + percents
  geom_text(
    aes(label = Label),
    hjust = -0.1, fontface = "bold", size = 3.4, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.15),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles & caption (consistent with Q22 format)
  labs(
    title = "Q8: AI Tools/ Platforms Used in Teaching",
    subtitle = "Multiple selections allowed; percentages based on valid faculty responses",
    x = "Percentage of Faculty",
    y = "AI Tool/ Platform",
    caption = paste0(
      "n = ", n_valid
    )
  ) +

  # Unified publication-style theme (same as Q22)
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.0, lineheight = 1.1),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin = margin(10, 65, 35, 15)
  )

# ---- 4) Save high-resolution plot ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q8_AI_Tools_Used_Teaching_horizontal_pub.png",
  plot = p_q8F,
  dpi = 600,
  width = 8.9,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot and summary table ----
print(p_q8F)
knitr::kable(q8_tab, caption = paste0("Q8: AI Tools/ Platforms Used in Teaching (n = ", n_valid, ")"))
```

## Q9: How would you rate your ability to perform the following tasks using AI tools in your teaching?


```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q9_cols <- names(faculty_analysis_data)[grepl("^q9[_. ]?[1-5]$", tolower(names(faculty_analysis_data)))]
q9_cols <- q9_cols[order(as.integer(str_extract(q9_cols, "\\d+$")))]
stopifnot(length(q9_cols) == 5)

resp_levels <- c("Expert","Proficient","Intermediate","Beginner","No Experience")
task_labels <- c(
  "Implement AI-assisted coding environments for teaching",
  "Use AI tools for real-time data analysis and visualization in class",
  "Automate student feedback and assessment using AI tools",
  "Guide students in applying AI to solve real-world data science problems",
  "Employ AI for student performance tracking & curriculum adjustments"
)

q9_long <- faculty_analysis_data %>%
  pivot_longer(all_of(q9_cols), names_to = "item", values_to = "Response", values_drop_na = FALSE) %>%
  mutate(
    idx  = as.integer(str_extract(item, "\\d+$")),
    Task = factor(task_labels[idx], levels = task_labels),
    # normalize any odd spacing and map to ordered factor
    Response = Response |> as.character() |> str_replace_all("\\u00A0", " ") |> str_squish(),
    Response = factor(Response, levels = resp_levels)
  ) %>%
  filter(!is.na(Task), !is.na(Response))

# ---- Summary ----
grid <- tidyr::expand_grid(
  Task = factor(task_labels, levels = task_labels),
  Response = factor(resp_levels, levels = resp_levels)
)

q9_tab <- grid %>%
  left_join(q9_long %>% count(Task, Response, name = "Count"), by = c("Task","Response")) %>%
  mutate(Count = coalesce(Count, 0L)) %>%
  group_by(Task) %>%
  mutate(Percent = round(100 * Count / pmax(sum(Count), 1), 1)) %>%
  ungroup()

knitr::kable(q9_tab, caption = "Q9: Ability to Perform Teaching Tasks Using AI Tools — Summary")


```


## Q9 fixed color theme

```{r}
# ---- 1) Define faculty Likert color palette ----
palette_likert <- c(
  "No Experience" = "#B2182B",  # dark red
  "Beginner"      = "#E67E22",  # orange
  "Intermediate"  = "#F1C40F",  # yellow
  "Proficient"    = "#48C9B0",  # teal
  "Expert"        = "#117A65"   # dark green
)

# ---- 2) Compute summary for plotting ----
plot_df_summary <- q9_long %>%
  count(Task, Response, name = "Count") %>%
  group_by(Task) %>%
  mutate(Percent = Count / sum(Count)) %>%
  ungroup() %>%
  mutate(
    Task = factor(stringr::str_wrap(as.character(Task), 34),
                  levels = stringr::str_wrap(task_labels, 34)),
    Response = factor(Response, levels = rev(resp_levels), ordered = TRUE),  # reversed so "Expert" at top legend
    Label = ifelse(Percent > 0.04, paste0(round(Percent * 100), "%"), "")
  )

# ---- 3) Horizontal 100% stacked bar chart ----
p_q9F <- ggplot(plot_df_summary, aes(y = Task, x = Percent, fill = Response)) +
  geom_col(color = "gray30", width = 0.7) +
  geom_text(
    aes(label = Label),
    position = position_stack(vjust = 0.5),
    color = "blue",
    size = 3,
    fontface = "bold"
  ) +

  # Scales
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(
    values = palette_likert,
    name = "Skill Level",
    guide = guide_legend(reverse = TRUE)
  ) +

  # Labels
  labs(
    title = "Q9: Ability to Perform Teaching Tasks \nUsing AI Tools",
    x = "Percentage of Faculty",
    y = "Teaching Task"
  ) +

  # Theme
  theme_publication() +
  theme(
    axis.text.y = element_text( size = 10.5, lineheight = 1.1),
    axis.text.x = element_text(face = "bold", size = 10.5),
    plot.title  = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 60, 35, 15)
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q9_faculty_ability_AI_tasks.png",
  plot = p_q9F,
  dpi = 700,
  width = 10.5,
  height = 6.5,
  units = "in"
)

# ---- 5) Display ----
print(p_q9F)

```


## Q10: How frequently do you use AI tools in the following teaching activities?


```{r Q10}
# ---- 1) Select dataset ----
faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# ---- 2) Identify columns ----
q10_cols <- names(faculty_analysis_data)[grepl("^q10[_. ]?[1-4]$", tolower(names(faculty_analysis_data)))]
if (length(q10_cols) == 0) {
  q10_cols <- names(faculty_analysis_data)[grepl("^q10(\\.|_|\\[|:)?\\d+$", tolower(names(faculty_analysis_data)))]
}
q10_cols <- q10_cols[order(as.integer(stringr::str_extract(q10_cols, "\\d+$")))]
stopifnot(length(q10_cols) == 4)

# ---- 3) Define levels ----
resp_levels <- c("Never","Rarely","Sometimes","Often","Always")
task_labels <- c(
  "Automating grading or feedback for data science assignments",
  "Engaging students through AI-assisted interactive labs (e.g., coding or machine learning exercises)",
  "Monitoring student learning progress using AI-based analytics",
  "Generating teaching materials with AI support"
)


# ---- 4) Reshape + robust mapping ----
q10_long <- faculty_analysis_data %>%
  pivot_longer(all_of(q10_cols), names_to = "item", values_to = "resp_raw", values_drop_na = FALSE) %>%
  mutate(
    idx  = as.integer(stringr::str_extract(item, "\\d+$")),
    Task = factor(task_labels[idx], levels = task_labels),
    txt  = as.character(resp_raw) |> stringr::str_replace_all("\\u00A0", " ") |> stringr::str_squish(),
    num  = suppressWarnings(as.numeric(txt)),
    Response = dplyr::case_when(
      !is.na(num) & num == 1 ~ "Never",
      !is.na(num) & num == 2 ~ "Rarely",
      !is.na(num) & num == 3 ~ "Sometimes",
      !is.na(num) & num == 4 ~ "Often",
      !is.na(num) & num == 5 ~ "Always",
      str_detect(txt, regex("^never$",     TRUE)) ~ "Never",
      str_detect(txt, regex("^rarely$",    TRUE)) ~ "Rarely",
      str_detect(txt, regex("^sometimes?$",TRUE)) ~ "Sometimes",
      str_detect(txt, regex("^often$",     TRUE)) ~ "Often",
      str_detect(txt, regex("^always$",    TRUE)) ~ "Always",
      TRUE ~ NA_character_
    ),
    Response = factor(Response, levels = resp_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(Task), !is.na(Response))

# ---- 5) Summary table ----
grid <- tidyr::expand_grid(
  Task = factor(task_labels, levels = task_labels),
  Response = factor(resp_levels, levels = resp_levels, ordered = TRUE)
)

q10_tab <- grid %>%
  left_join(q10_long %>% count(Task, Response, name = "Count"), by = c("Task","Response")) %>%
  mutate(Count = dplyr::coalesce(Count, 0L)) %>%
  group_by(Task) %>%
  mutate(Percent = round(100 * Count / pmax(sum(Count), 1), 1)) %>%
  ungroup()

knitr::kable(q10_tab, caption = "Q10: Frequency of Using AI Tools in Teaching Activities — Summary")

# ---- 1) Define custom color palette (faculty Likert) ----
palette_likert <- c(
  "Never" = "#B2182B",      # dark red
  "Rarely" = "#E67E22",     # orange
  "Sometimes" = "#F1C40F",  # yellow
  "Often" = "#48C9B0",      # teal
  "Always" = "#117A65"      # dark green
)

# ---- 2) Compute summary for plotting ----
plot_df_summary <- q10_long %>%
  count(Task, Response, name = "Count") %>%
  group_by(Task) %>%
  mutate(Percent = Count / sum(Count)) %>%
  ungroup() %>%
  mutate(
    Task = factor(stringr::str_wrap(as.character(Task), 34),
                  levels = stringr::str_wrap(task_labels, 34)),
    Response = factor(Response, levels = resp_levels, ordered = TRUE),
    Label = ifelse(Percent > 0.04, paste0(round(Percent * 100), "%"), "")
  )

# ---- 3) Horizontal 100% stacked bar chart ----
p_q10F <- ggplot(plot_df_summary, aes(y = Task, x = Percent, fill = Response)) +
  geom_col(color = "gray30", width = 0.7) +

  # Inline percent labels
  geom_text(
    aes(label = Label),
    position = position_stack(vjust = 0.5),
    color = "blue",
    size = 3,
    fontface = "bold"
  ) +

  # Scales
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(
    values = palette_likert,
    name = "Response",
    guide = guide_legend(reverse = TRUE)
  ) +

  # Labels
  labs(
    title = "Q10: Frequency of Using AI Tools in \nTeaching Activities",
    x = "Percentage of Responses",
    y = "Teaching Task"
  ) +

  # Theme
  theme_publication() +
  theme(
    axis.text.y = element_text(face = "bold", size = 10, lineheight = 1.1),
    axis.text.x = element_text(face = "bold", size = 10),
    plot.title  = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 60, 35, 15)
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q10_faculty_frequency_AI_teaching.png",
  plot = p_q10F,
  dpi = 700,
  width = 9.5,
  height = 6.5,
  units = "in"
)

# ---- 5) Display ----
print(p_q10F)
```


## Q11

```{r}
# ---- 1) Data selection ----
faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q11_cols <- names(faculty_analysis_data)[grepl("^q11[_. ]?[1-5]$", tolower(names(faculty_analysis_data)))]
if (length(q11_cols) == 0) {
  q11_cols <- names(faculty_analysis_data)[grepl("^q11(\\.|_|\\[|:)?\\d+$", tolower(names(faculty_analysis_data)))]
}
q11_cols <- q11_cols[order(as.integer(stringr::str_extract(q11_cols, "\\d+$")))]
stopifnot(length(q11_cols) == 5)

resp_levels <- c("Strongly Disagree","Disagree","Neutral","Agree","Strongly Agree")
item_labels <- c(
  "AI tools are vital for \nfuture data science education.",
  "AI can significantly improve \nthe teaching of data science topics like machine learning, deep learning, and big data.",
  "Over-reliance on AI tools may \nreduce the development of critical thinking skills in students.",
  "AI tools can efficiently automate \nrepetitive teaching tasks, freeing up time for more complex educational engagements.",
  "AI tools are capable of improving \nequity and inclusion in data science education."
)

# ---- 2) Long format ----
q11_long <- faculty_analysis_data %>%
  pivot_longer(all_of(q11_cols), names_to = "item", values_to = "resp_raw", values_drop_na = FALSE) %>%
  mutate(
    idx  = as.integer(stringr::str_extract(item, "\\d+$")),
    Statement = factor(item_labels[idx], levels = item_labels),
    txt  = str_squish(str_replace_all(as.character(resp_raw), "\\u00A0", " ")),
    num  = suppressWarnings(as.numeric(txt)),
    Response = case_when(
      num == 1 ~ "Strongly Disagree",
      num == 2 ~ "Disagree",
      num == 3 ~ "Neutral",
      num == 4 ~ "Agree",
      num == 5 ~ "Strongly Agree",
      str_detect(txt, regex("^strongly\\s*disagree$", TRUE)) ~ "Strongly Disagree",
      str_detect(txt, regex("^disagree$", TRUE))              ~ "Disagree",
      str_detect(txt, regex("^neutral$", TRUE))               ~ "Neutral",
      str_detect(txt, regex("^agree$", TRUE))                 ~ "Agree",
      str_detect(txt, regex("^strongly\\s*agree$", TRUE))     ~ "Strongly Agree",
      TRUE ~ NA_character_
    ),
    Response = factor(Response, levels = resp_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(Statement), !is.na(Response))

# ---- 3) Plot dataframe ----
plot_data <- q11_long %>%
  mutate(
    Statement = factor(
      str_wrap(as.character(Statement), 40),
      levels = str_wrap(item_labels, 40)
    ),
    Response = factor(Response, levels = resp_levels, ordered = TRUE)
  )

# ---- 4) Horizontal 100% stacked bar chart ----
# ---- 1) Define Likert color palette (same as student version) ----
palette_likert <- c(
  "Strongly Disagree" = "#B2182B",  # dark red
  "Disagree"          = "#E67E22",  # orange
  "Neutral"           = "#F1C40F",  # yellow
  "Agree"             = "#48C9B0",  # teal
  "Strongly Agree"    = "#117A65"   # dark green
)

# ---- 2) Compute summary for plotting ----
plot_df_summary <- q11_long %>%
  count(Statement, Response, name = "Count") %>%
  group_by(Statement) %>%
  mutate(Percent = Count / sum(Count)) %>%
  ungroup() %>%
  mutate(
    Statement = factor(
      stringr::str_wrap(as.character(Statement), 40),
      levels = stringr::str_wrap(item_labels, 40)
    ),
    Response = factor(Response, levels = resp_levels, ordered = TRUE),
    Label = ifelse(Percent > 0.04, paste0(round(Percent * 100), "%"), "")
  )

# ---- 3) Horizontal 100% stacked bar chart ----
p_q11F <- ggplot(plot_df_summary, aes(y = Statement, x = Percent, fill = Response)) +
  geom_col(color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = Label),
    position = position_stack(vjust = 0.5),
    color = "black",
    size = 3,
    fontface = "bold"
  ) +

  # Scales and colors
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(
    values = palette_likert,
    name = "Response",
    guide = guide_legend(reverse = TRUE)
  ) +

  # Titles and labels
  labs(
    title = "Q11: Faculty Attitudes Toward AI in Data Science Education",
    x = "Percentage of Responses",
    y = "Statement"
  ) +

  # Publication-style theme
  theme_publication() +
  theme(
    axis.text.y = element_text(face = "bold", size = 10.5, lineheight = 1.1),
    axis.text.x = element_text(face = "bold", size = 10.5),
    plot.title  = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 60, 35, 15)
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q11_faculty_attitudes_AI_education.png",
  plot = p_q11F,
  dpi = 700,
  width = 9.5,
  height = 6.5,
  units = "in"
)

# ---- 5) Display ----
print(p_q11F)

```


## Q12: Do you believe that AI tools should complement or replace traditional methods of teaching data science?
```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q12_col <- dplyr::case_when(
  "Q12" %in% names(faculty_analysis_data) ~ "Q12",
  "q12" %in% names(faculty_analysis_data) ~ "q12",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q12_col))

# ---- Canonical labels in instrument order ----
levels_q12 <- c(
  "AI tools should complement traditional methods.",
  "AI tools should balance with traditional methods.",
  "AI tools should replace traditional methods in some contexts.",
  "AI tools should not replace traditional methods."
)

# ---- Map numeric or text to canonical labels ----
faculty_analysis_data <- faculty_analysis_data %>%
  mutate(
    q12_raw = .data[[q12_col]],
    txt = as.character(q12_raw) |>
      str_replace_all("\\u00A0", " ") |>
      str_squish(),
    num = suppressWarnings(as.numeric(txt)),
    q12_std = case_when(
      !is.na(num) & num == 1 ~ levels_q12[1],
      !is.na(num) & num == 2 ~ levels_q12[2],
      !is.na(num) & num == 3 ~ levels_q12[3],
      !is.na(num) & num == 4 ~ levels_q12[4],
      str_detect(txt, regex("\\bcomplement\\b",  TRUE)) ~ levels_q12[1],
      str_detect(txt, regex("\\bbalance\\b",     TRUE)) ~ levels_q12[2],
      str_detect(txt, regex("replace.*some",     TRUE)) ~ levels_q12[3],
      str_detect(txt, regex("not\\s*replace",    TRUE)) ~ levels_q12[4],
      TRUE ~ NA_character_
    ) |> factor(levels = levels_q12)
  )


# ---- 1) Summary table ----
q12_tab <- faculty_analysis_data %>%
  filter(!is.na(q12_std) & q12_std != "") %>%   # keep valid responses only
  count(q12_std, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  )

n_valid <- sum(q12_tab$Count)

# ---- 2) Prepare plotting dataframe with wrapped labels ----
plot_data <- q12_tab %>%
  mutate(option = factor(
    str_wrap(as.character(q12_std), 38),
    levels = str_wrap(levels_q12, 38)
  ))

# ---- 3) Horizontal bar chart with refined publication theme ----
p_q12 <- ggplot(plot_data, aes(
  y = fct_reorder(option, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline count + percent labels
  geom_text(
    aes(label = Label),
    hjust = -0.1, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting (percentage scale)
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and captions
  labs(
    title    = "Q12: Complement vs. Replace Traditional Methods",
#    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Response",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Consistent publication theme (same as Q22)
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 60, 35, 15)   # extra right margin for labels
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q12_Complement_vs_Replace_horizontal_pub.png",
  plot = p_q12,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot and summary table ----
print(p_q12)
knitr::kable(q12_tab, caption = paste0("Q12: Complement vs. Replace (n = ", n_valid, ")"))
```


## Q13: How concerned are you about the following aspects of AI integration in data science education?


```{r}
# ---- 0) Data source ----
faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# ---- 1) Detect Q13 sub-items ----
q13_cols <- names(faculty_analysis_data)[grepl("^q13[_. ]?[1-5]$", tolower(names(faculty_analysis_data)))]
if (length(q13_cols) == 0) {
  q13_cols <- names(faculty_analysis_data)[grepl("^q13(\\.|_|\\[|:)?\\d+$", tolower(names(faculty_analysis_data)))]
}
q13_cols <- q13_cols[order(as.integer(str_extract(q13_cols, "\\d+$")))]
stopifnot(length(q13_cols) == 5)

# ---- 2) Define levels, labels, and palette ----
resp_levels <- c("Very Concerned","Somewhat Concerned","Neutral","Not Concerned")

item_labels <- c(
  "AI may produce biased results",
  "Lack of transparency in how AI generates outputs",
  "Keeping up with evolving AI technologies",
  "Overreliance on AI tools \nreducing faculty–student engagement",
  "Data privacy and security issues"
)


# ---- Red–Teal publication palette ----
pal_q13 <- c(
  "Very Concerned"     =  "#B2182B",
  "Somewhat Concerned" =  "#E67E22",
  "Neutral"            =  "#F1C40F",
  "Not Concerned"      =  "#48C9B0"
)

# ---- 3) Reshape and clean ----
q13_long <- faculty_analysis_data %>%
  pivot_longer(all_of(q13_cols), names_to = "item", values_to = "resp_raw", values_drop_na = FALSE) %>%
  mutate(
    idx  = as.integer(str_extract(item, "\\d+$")),
    Concern = factor(item_labels[idx], levels = item_labels),
    txt  = str_replace_all(as.character(resp_raw), "\\u00A0", " ") |> str_squish(),
    num  = suppressWarnings(as.numeric(txt)),
    Response = case_when(
      !is.na(num) & num == 1 ~ "Very Concerned",
      !is.na(num) & num == 2 ~ "Somewhat Concerned",
      !is.na(num) & num == 3 ~ "Neutral",
      !is.na(num) & num == 4 ~ "Not Concerned",
      str_detect(txt, regex("^very\\s*concerned$", TRUE))     ~ "Very Concerned",
      str_detect(txt, regex("^somewhat\\s*concerned$", TRUE)) ~ "Somewhat Concerned",
      str_detect(txt, regex("^neutral$", TRUE))               ~ "Neutral",
      str_detect(txt, regex("^not\\s*concerned$", TRUE))      ~ "Not Concerned",
      TRUE ~ NA_character_
    ),
    Response = factor(Response, levels = resp_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(Concern), !is.na(Response))

# ---- 4) Calculate proportions ----
plot_df_q13 <- q13_long %>%
  count(Concern, Response, .drop = FALSE) %>%
  group_by(Concern) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Concern = factor(str_wrap(Concern, 45),
                     levels = str_wrap(item_labels, 45))
  )

# ---- 5) Horizontal stacked bar with text labels ----
p_fac_q13 <- ggplot(plot_df_q13, aes(y = Concern, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = pal_q13, guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Faculty Concerns About AI Integration \ninto Data Science Education",
  #  subtitle = "Proportion of faculty indicating level of concern for each issue",
    x = "Percentage of Responses",
    y = "Concern",
    fill = "Response",
 #   caption = "Note: Percentages are calculated of faculty responses within each concern."
  ) +
  theme_publication()

# ---- 6) Save high-resolution PNG ----
ggsave(
  filename = "Faculty_figures/Q13_faculty_concerns.png",
  plot = p_fac_q13,
  dpi = 700,
  width = 9.5,
  height = 6.5,
  units = "in"
)

print(p_fac_q13)

```


##  Q14: What are the biggest challenges you face in integrating AI tools into your data science courses?

```{r}
faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# ---- Canonical labels in instrument order ----
opt_map <- c(
  "1" = "Difficulty in learning \nand mastering AI tools",
  "2" = "Insufficient institutional \nsupport or training",
  "3" = "Ethical concerns \n(fairness, bias)",
  "4" = "Limited access to advanced \nAI tools or platforms",
  "5" = "Resistance from students \nor faculty to adopt AI tools",
  "6" = "Balancing AI tools with \nthe curriculum’s learning objectives"
)
canon_levels <- unname(opt_map)

# # # ---- Detect export style ----
nm <- names(faculty_analysis_data)
q14_single <- nm[tolower(nm) == "q14"]
q14_wide   <- nm[str_detect(tolower(nm), "^q14(\\.|_|\\s|\\[|:)?\\d+$")]
q14_wide   <- setdiff(q14_wide, nm[str_detect(tolower(nm), "text|comment")])

faculty_analysis_data <- mutate(faculty_analysis_data, .row = row_number())

# ---- Tidy to long format (rows = respondent selections) ----
if (length(q14_wide) >= 1) {
  # A) Multi-column export Q14_1..Q14_6
  long <- faculty_analysis_data %>%
    select(.row, all_of(q14_wide)) %>%
    pivot_longer(-.row, names_to = "col", values_to = "val") %>%
    mutate(
      idx = str_extract(col, "\\d+$"),
      option = unname(opt_map[idx]),
      val_chr = tolower(str_squish(as.character(val))),
      selected = case_when(
        val_chr %in% c("1","true","t","yes","y","selected","checked") ~ TRUE,
        val_chr %in% c("0","false","f","no","n","not selected","")     ~ FALSE,
        suppressWarnings(!is.na(as.numeric(val_chr)) & as.numeric(val_chr) > 0) ~ TRUE,
        TRUE ~ NA
      )
    ) %>%
    filter(!is.na(option), isTRUE(selected)) %>%
    select(.row, option)
} else if (length(q14_single) == 1) {
  # B) Single column with comma/semicolon separated choices
  long <- faculty_analysis_data %>%
    select(.row, q14 = all_of(q14_single)) %>%
    filter(!is.na(q14), q14 != "") %>%
    mutate(q14 = str_split(q14, "[,;]")) %>%
    unnest(q14) %>%
    mutate(q14 = str_squish(q14),
           option = case_when(
             str_detect(q14, regex("difficulty|master", TRUE))         ~ opt_map["1"],
             str_detect(q14, regex("support|training", TRUE))          ~ opt_map["2"],
             str_detect(q14, regex("ethic|fairness|bias", TRUE))       ~ opt_map["3"],
             str_detect(q14, regex("access|platform", TRUE))           ~ opt_map["4"],
             str_detect(q14, regex("resistance|adopt", TRUE))          ~ opt_map["5"],
             str_detect(q14, regex("curriculum|objective", TRUE))      ~ opt_map["6"],
             TRUE ~ NA_character_
           )) %>%
    filter(!is.na(option)) %>%
    select(.row, option)
} else {
  stop("Q14 columns not found.")
}
# #
# ---- Summary table ----
denom <- length(unique(long$.row))

```


## (redone no duplicates)

```{r}
# --- Sanity checks ---
denom <- n_distinct(long$.row)   # number of valid (non-empty) respondents
total_mentions <- nrow(long)     # total selections across all respondents
message("Respondents (denom): ", denom, " | Total selections (mentions): ", total_mentions)


# --- Safety: ensure no duplicate (.row, option) pairs are counted twice ---
long_nodup <- long %>% distinct(.row, option)

# --- Summary table (counts & percents based on respondents) ---
q14_tab <- long_nodup %>%
  count(option, name = "Count") %>%
  complete(option = canon_levels, fill = list(Count = 0)) %>%
  mutate(
    Percent = Count / denom,
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Count))  # sort by frequency

# optional: print table
knitr::kable(
  q14_tab %>% rename(Option = option),
  caption = paste0("Q14: Biggest Challenges Integrating AI Tools — Summary (n = ", denom, ")")
)

# --- Plot (Q22 style: horizontal, counts+percents, same colors/fonts) ---
p_q14 <- ggplot(q14_tab, aes(
  y = fct_reorder(option, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +
  geom_text(aes(label = Label),
            hjust = -0.1, fontface = "bold", size = 3.6, color = "black") +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(q14_tab$Percent) + 0.15),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title    = "Q14: Challenges Integrating AI Tools in Teaching",
    subtitle = "Multiple selections allowed; percentages based on valid faculty respondents",
    x        = "Percentage of Faculty",
    y        = "Challenge",
    caption  = paste0(
      "n = ", denom,
      " faculty respondents."
    )
  ) +
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 60, 35, 15)   # extra right margin for labels
  )


ggsave(
  filename = "Faculty_figures/Q14_Challenges_AI.png",
  plot = p_q14,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot and summary table ----
print(p_q14)
knitr::kable(
  q14_tab %>% rename(Option = option),
  caption = paste0("Q14: Biggest Challenges Integrating AI Tools (n = ", n_valid, ")")
)
```

## Q15: To what extent do you agree that AI tools can enhance the following aspects of data science teaching?

```{r, warning=FALSE, message=FALSE}
# --- Faculty Q15: Agreement that AI Tools Enhance Aspects of DS Teaching ---

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# Detect relevant Q15 columns
q15_cols <- names(faculty_analysis_data)[grepl("^q15[_. ]?[1-5]$", tolower(names(faculty_analysis_data)))]
if (length(q15_cols) == 0) {
  q15_cols <- names(faculty_analysis_data)[grepl("^q15(\\.|_|\\[|:)?\\d+$", tolower(names(faculty_analysis_data)))]
}
q15_cols <- q15_cols[order(as.integer(stringr::str_extract(q15_cols, "\\d+$")))]
stopifnot(length(q15_cols) == 5)

# --- Response levels and improved line-broken item labels ---
resp_levels <- c("Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree")

item_labels <- c(
  "Automating assessment \nand feedback for DS projects",
  "Personalizing learning paths \nusing AI-driven insights",
  "Simplifying complex DS concepts \nwith AI-enhanced visualizations",
  "Supporting real-time tracking \nand identifying student challenges",
  "Enhancing engagement via \nAI-driven interactive modules"
)

# --- Map responses numerically or textually ---
q15_long <- faculty_analysis_data %>%
  pivot_longer(all_of(q15_cols), names_to = "item", values_to = "resp_raw", values_drop_na = FALSE) %>%
  mutate(
    idx  = as.integer(stringr::str_extract(item, "\\d+$")),
    Aspect = factor(item_labels[idx], levels = item_labels),
    txt  = as.character(resp_raw) |> stringr::str_replace_all("\\u00A0", " ") |> stringr::str_squish(),
    num  = suppressWarnings(as.numeric(txt)),
    Response = dplyr::case_when(
      num == 1 ~ "Strongly Disagree",
      num == 2 ~ "Disagree",
      num == 3 ~ "Neutral",
      num == 4 ~ "Agree",
      num == 5 ~ "Strongly Agree",
      str_detect(txt, regex("^strongly\\s*disagree$", TRUE)) ~ "Strongly Disagree",
      str_detect(txt, regex("^disagree$", TRUE))              ~ "Disagree",
      str_detect(txt, regex("^neutral$", TRUE))               ~ "Neutral",
      str_detect(txt, regex("^agree$", TRUE))                 ~ "Agree",
      str_detect(txt, regex("^strongly\\s*agree$", TRUE))     ~ "Strongly Agree",
      TRUE ~ NA_character_
    ),
    Response = factor(Response, levels = resp_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(Aspect), !is.na(Response))

# --- Compute proportions ---
plot_data <- q15_long %>%
  count(Aspect, Response, .drop = FALSE) %>%
  group_by(Aspect) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  group_by(Aspect) %>%
  mutate(pct_agree = sum(pct[Response %in% c("Agree","Strongly Agree")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Aspect = fct_reorder(Aspect, pct_agree, .desc = TRUE))

# --- Define publication red–teal Likert palette ---
pal_q15 <- c(
  "Strongly Disagree" = "#B2182B",
  "Disagree"          = "#E67E22",
  "Neutral"           = "#F1C40F",
  "Agree"             = "#48C9B0",
  "Strongly Agree"    = "#117A65"
)

# --- Plot ---
# --- Plot with increased vertical spacing between bars ---
p_q15_faculty <- ggplot(plot_data, aes(y = Aspect, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.55) +  # Reduced width adds vertical space
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold", lineheight = 0.9
  ) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = pal_q15, guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Q15: Faculty Opinions on AI Tools' \nPotential to Enhance DS Teaching",
  #  subtitle = "Responses grouped by agreement level",
    x = "Percentage of Responses",
    y = "Teaching Enhancement Aspect",
    fill = "Response",
  #  caption = "Note: Percentages represent the proportion of responses within each aspect category."
  ) +
  theme_publication() +
  theme(
    axis.text.y = element_text(face = "bold", lineheight = 1.1),  # improves readability
    panel.spacing.y = grid::unit(0.5, "lines"),                         # adds vertical padding between facets (if used)
    plot.margin = margin(10, 20, 45, 20)                          # a bit more bottom space for legend
  )

# --- Save High-Resolution Figure ---
ggsave(
  filename = "Faculty_figures/Q15_AI_DS_Teaching.png",
  plot = p_q15_faculty,
  dpi = 700,
  width = 9.5,
  height = 6.8,  # slightly taller for improved readability
  units = "in"
)

print(p_q15_faculty)


```


## Q17: How aware are you of your institution's guidelines on the use of AI in teaching and assessments?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q17_col <- dplyr::case_when(
  "Q17" %in% names(faculty_analysis_data) ~ "Q17",
  "q17" %in% names(faculty_analysis_data) ~ "q17",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q17_col))

# Map to instrument labels
aware_levels <- c("Very Aware", "Somewhat Aware", "Neutral", "Not Aware")

dat_q17 <- faculty_analysis_data %>%
  mutate(
    q17_raw = .data[[q17_col]],
    q17_txt = str_squish(as.character(q17_raw)),
    q17_num = suppressWarnings(as.numeric(q17_txt)),
    q17_std = case_when(
      !is.na(q17_num) & q17_num == 1 ~ "Very Aware",
      !is.na(q17_num) & q17_num == 2 ~ "Somewhat Aware",
      !is.na(q17_num) & q17_num == 3 ~ "Neutral",
      !is.na(q17_num) & q17_num == 4 ~ "Not Aware",
      str_detect(q17_txt, regex("^very\\s*aware$", TRUE))     ~ "Very Aware",
      str_detect(q17_txt, regex("^somewhat\\s*aware$", TRUE)) ~ "Somewhat Aware",
      str_detect(q17_txt, regex("^neutral$", TRUE))           ~ "Neutral",
      str_detect(q17_txt, regex("^not\\s*aware$", TRUE))      ~ "Not Aware",
      TRUE ~ NA_character_
    ) |> factor(levels = aware_levels, ordered = TRUE)
  )

dat_q17_clean <- dat_q17 %>% filter(!is.na(q17_std))


# ---- 1) Compute Counts, Percents, and Labels ----
q17_tab <- dat_q17_clean  %>%
  filter(!is.na(q17_std) & q17_std != "") %>%
  count(q17_std, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Percent))

n_valid <- sum(q17_tab$Count)

# ---- 2) Prepare data for plotting ----
plot_data <- q17_tab %>%
  mutate(Response = str_wrap(as.character(q17_std), 44))

# ---- 3) Horizontal Bar Chart ----
p_q17_faculty <- ggplot(plot_data, aes(
  y = fct_reorder(Response, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels (Count + Percent)
  geom_text(
    aes(label = Label),
    hjust = -0.05,
    fontface = "bold",
    size = 3.6,
    color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles & caption
  labs(
    title    = "Q17: Awareness of Institutional AI Guidelines",
    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Response",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Publication-style theme (consistent with Q18–Q23)
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1, face = "bold"),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 100, 35, 15)  # add right padding for labels
  )

# ---- 4) Save High-Resolution PNG ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q17_faculty_awareness_AI_guidelines_horizontal_pub.png",
  plot = p_q17_faculty,
  dpi = 700,
  width = 9.0,
  height = 5.5,
  units = "in"
)

# ---- 5) Display Plot and Table ----
print(p_q17_faculty)

knitr::kable(
  q17_tab %>% rename(Response = q17_std),
  caption = paste0("Q17: Awareness of Institutional AI Guidelines (n = ", n_valid, ")")
)
```


## Q18: What primary ethical concern do you have about using AI tools in education?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

nm <- names(faculty_analysis_data)
q18_col <- dplyr::case_when(
  "Q18" %in% nm ~ "Q18",
  "q18" %in% nm ~ "q18",
  TRUE ~ {
    cands <- nm[grepl("^q18", tolower(nm))]
    cands <- cands[!grepl("text|comment|other", tolower(cands))]
    if (length(cands)) cands[1] else NA_character_
  }
)
stopifnot(!is.na(q18_col))

eth_levels <- c(
  "Bias in AI-driven student evaluations",
  "Infringement on student privacy with AI monitoring tools",
  "Overreliance on AI leading to less human oversight in teaching and assessments"
)

# ---- Standardize responses (numeric or text) ----
df_q18 <- faculty_analysis_data %>%
  mutate(
    q18_raw = .data[[q18_col]],
    txt  = str_squish(as.character(q18_raw)),
    num  = suppressWarnings(as.numeric(txt)),
    q18_std = case_when(
      !is.na(num) & num == 1 ~ eth_levels[1],
      !is.na(num) & num == 2 ~ eth_levels[2],
      !is.na(num) & num == 3 ~ eth_levels[3],
      str_detect(txt, regex("bias|evaluation", TRUE))             ~ eth_levels[1],
      str_detect(txt, regex("privacy|monitor", TRUE))             ~ eth_levels[2],
      str_detect(txt, regex("over[- ]?reliance|oversight", TRUE)) ~ eth_levels[3],
      TRUE ~ NA_character_
    ) |> factor(levels = eth_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(q18_std))


# ---- 1) Summary table ----
q18_tab <- df_q18 %>%
  filter(!is.na(q18_std) & q18_std != "") %>%   # keep valid responses
  count(q18_std, name = "Count") %>%
  complete(q18_std = factor(eth_levels, levels = eth_levels, ordered = TRUE),
           fill = list(Count = 0)) %>%
  mutate(
    Percent = Count / sum(Count),
    Label = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Count))  # sort in decreasing order

n_valid <- sum(q18_tab$Count)

# ---- 2) Wrap long labels ----
wrap_w <- 40
wrapped_levels <- stringr::str_wrap(eth_levels, wrap_w)
plot_data <- q18_tab %>%
  mutate(Response_w = stringr::str_wrap(as.character(q18_std), wrap_w))

# ---- 3) Horizontal bar chart (publication style) ----
p_q18 <- ggplot(plot_data, aes(
  y = fct_reorder(Response_w, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline count + percent labels
  geom_text(
    aes(label = Label),
    hjust = -0.1, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis percentage formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.15),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles & captions
  labs(
    title    = "Q18: Primary Ethical Concern in Using AI Tools",
    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Ethical Concern",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Apply publication theme (same as Q22, Q12)
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 60, 35, 15)   # extra right margin for labels
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q18_Primary_Ethical_Concern_horizontal_pub.png",
  plot = p_q18,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot and summary table ----
print(p_q18)
knitr::kable(
  q18_tab %>% rename(Response = q18_std),
  caption = paste0("Q18: Primary Ethical Concern (n = ", n_valid, ")")
)
```


## Q19: How important is it to maintain human oversight when using AI for assessments and student feedback?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

nm <- names(faculty_analysis_data)
q19_col <- dplyr::case_when(
  "Q19" %in% nm ~ "Q19",
  "q19" %in% nm ~ "q19",
  TRUE ~ {
    cands <- nm[grepl("^q19", tolower(nm))]
    cands <- cands[!grepl("text|comment|other", tolower(cands))]
    if (length(cands)) cands[1] else NA_character_
  }
)
stopifnot(!is.na(q19_col))

imp_levels <- c("Very Important", "Important", "Neutral", "Not Important")

df_q19 <- faculty_analysis_data %>%
  mutate(
    q19_raw = .data[[q19_col]],
    txt = str_squish(as.character(q19_raw)),
    num = suppressWarnings(as.numeric(txt)),
    q19_std = case_when(
      !is.na(num) & num == 1 ~ imp_levels[1],
      !is.na(num) & num == 2 ~ imp_levels[2],
      !is.na(num) & num == 3 ~ imp_levels[3],
      !is.na(num) & num == 4 ~ imp_levels[4],
      str_detect(txt, regex("^very\\s*important$", TRUE)) ~ imp_levels[1],
      str_detect(txt, regex("^important$", TRUE))         ~ imp_levels[2],
      str_detect(txt, regex("^neutral$", TRUE))           ~ imp_levels[3],
      str_detect(txt, regex("^not\\s*important$", TRUE))  ~ imp_levels[4],
      TRUE ~ NA_character_
    ) |> factor(levels = imp_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(q19_std))


```


## Q19 horizontal counts & percents
```{r}
# ---- 1) Compute counts & percents ----
q19_tab <- df_q19%>%
  filter(!is.na(q19_std) & q19_std != "") %>%
  count(q19_std, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Percent))   # sort by decreasing frequency

n_valid <- sum(q19_tab$Count)

# ---- 2) Prepare plot data ----
plot_data <- q19_tab %>%
  mutate(Response = str_wrap(as.character(q19_std), 40))

# ---- 3) Horizontal bar chart (publication style) ----
p_q19 <- ggplot(plot_data, aes(
  y = fct_reorder(Response, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels (counts + percents)
  geom_text(
    aes(label = Label),
    hjust = -0.05, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles & captions
  labs(
    title    = "Q19: Importance of Maintaining Human Oversight in AI-Assisted Teaching",
    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Response",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Theme consistent with other faculty figures
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 100, 35, 15)   # extra right margin for labels
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q19_Human_Oversight_horizontal_pub.png",
  plot = p_q19,
  dpi = 600,
  width = 8.8,
  height = 5.5,
  units = "in"
)

# ---- 5) Display results ----
print(p_q19)
knitr::kable(
  q19_tab %>% rename(Response = q19_std),
  caption = paste0("Q19: Importance of Maintaining Human Oversight (n = ", n_valid, ")")
)

```

## Q20: Which of the following best describe the AI policy in your syllabus/ class?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

# ---- Locate Q20 column (ignore text/comment fields) ----
nm <- names(faculty_analysis_data)
q20_col <- dplyr::case_when(
  "Q20" %in% nm ~ "Q20",
  "q20" %in% nm ~ "q20",
  TRUE ~ {
    cands <- nm[grepl("^q20", tolower(nm))]
    cands <- cands[!grepl("text|comment|other", tolower(cands))]
    if (length(cands)) cands[1] else NA_character_
  }
)
stopifnot(!is.na(q20_col))

# ---- Canonical labels in instrument order ----
policy_levels <- c(
  "AI is NOT allowed \nfor any activities/assignments",
  "Allowed for specific activities \ndesignated by instructor (disclosure required)",
  "Permitted for all \nactivities/assignments with proper attribution"
)

# ---- Standardize responses; ----
df_q20 <- faculty_analysis_data %>%
  mutate(
    q20_raw = .data[[q20_col]],
    txt = str_squish(as.character(q20_raw)),
    num = suppressWarnings(as.numeric(txt)),
    q20_std = case_when(
      # numeric exports
      !is.na(num) & num == 1 ~ policy_levels[1],
      !is.na(num) & num == 2 ~ policy_levels[2],
      !is.na(num) & num == 3 ~ policy_levels[3],
      # text exports (robust patterns)
      str_detect(txt, regex("not\\s+allowed|prohibit|not\\s+permitted", TRUE)) ~ policy_levels[1],
      str_detect(txt, regex("allowed\\s+for\\s+specific|specific\\s+activit|designat", TRUE)) ~ policy_levels[2],
      str_detect(txt, regex("permitted|allowed\\s+for\\s+all|proper\\s+attribution", TRUE)) ~ policy_levels[3],
      TRUE ~ NA_character_
    ) |> factor(levels = policy_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(q20_std))


# ---- 1) Summary table ----
q20_tab <- df_q20 %>%
  filter(!is.na(q20_std) & q20_std != "") %>%   # keep valid responses only
  count(q20_std, name = "Count") %>%
  complete(q20_std = factor(policy_levels, levels = policy_levels, ordered = TRUE),
           fill = list(Count = 0)) %>%
  mutate(
    Percent = Count / sum(Count),
    Label = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Count))   # sort by frequency

n_valid <- sum(q20_tab$Count)

# ---- 2) Wrap long labels ----
wrap_w <- 44
wrapped_levels <- stringr::str_wrap(policy_levels, wrap_w)
plot_data <- q20_tab %>%
  mutate(Response_w = stringr::str_wrap(as.character(q20_std), wrap_w))

# ---- 3) Horizontal bar chart (publication style) ----
p_q20 <- ggplot(plot_data, aes(
  y = fct_reorder(Response_w, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline count + percent labels
  geom_text(
    aes(label = Label),
    hjust = -0.05, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.15),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles & captions
  labs(
    title    = "Q20: AI Policy in Syllabus/Class",
    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Response",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Publication theme consistent with other figures
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 100, 35, 15)   # wider right margin for labels
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q20_AI_Policy_Syllabus_horizontal_pub.png",
  plot = p_q20,
  dpi = 600,
  width = 8.8,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot and table ----
print(p_q20)
knitr::kable(
  q20_tab %>% rename(Response = q20_std),
  caption = paste0("Q20: AI Policy in Syllabus/Class (n = ", n_valid, ")")
)
```


## Q21: How confident are you in your knowledge of AI tools available for data science education?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

nm <- names(faculty_analysis_data)
q21_col <- dplyr::case_when(
  "Q21" %in% nm ~ "Q21",
  "q21" %in% nm ~ "q21",
  TRUE ~ {
    cands <- nm[grepl("^q21", tolower(nm))]
    cands <- cands[!grepl("text|comment|other", tolower(cands))]
    if (length(cands)) cands[1] else NA_character_
  }
)
stopifnot(!is.na(q21_col))


conf_levels <- c("Very Confident", "Somewhat Confident", "Neutral", "Not Confident")

# ---- Standardize responses;  ----
df_q21 <- faculty_analysis_data %>%
  mutate(
    q21_raw = .data[[q21_col]],
    txt = str_squish(as.character(q21_raw)),
    num = suppressWarnings(as.numeric(txt)),
    q21_std = case_when(
      !is.na(num) & num == 1 ~ conf_levels[1],
      !is.na(num) & num == 2 ~ conf_levels[2],
      !is.na(num) & num == 3 ~ conf_levels[3],
      !is.na(num) & num == 4 ~ conf_levels[4],
      str_detect(txt, regex("^very\\s*confident$", TRUE))      ~ conf_levels[1],
      str_detect(txt, regex("^somewhat\\s*confident$", TRUE))  ~ conf_levels[2],
      str_detect(txt, regex("^neutral$", TRUE))                ~ conf_levels[3],
      str_detect(txt, regex("^not\\s*confident$", TRUE))       ~ conf_levels[4],
      TRUE ~ NA_character_
    ) |> factor(levels = conf_levels, ordered = TRUE)
  ) %>%
  filter(!is.na(q21_std))

# ---- Summary table  ----
q21_tab <- df_q21 %>%
  count(q21_std, name = "Count") %>%
  complete(q21_std = factor(conf_levels, levels = conf_levels, ordered = TRUE),
           fill = list(Count = 0)) %>%
  mutate(Percent = round(100 * Count / pmax(sum(Count), 1), 1)) %>%
  arrange(q21_std)

knitr::kable(q21_tab %>% rename(Response = q21_std),
             caption = "Q21: Confidence in Knowledge of AI Tools — Summary")

# ---- 1) Compute counts, percents, and labels ----
q21_tab <- df_q21 %>%
  filter(!is.na(q21_std) & q21_std != "") %>%
  count(q21_std, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Percent))

n_valid <- sum(q21_tab$Count)

# ---- 2) Prepare data for plotting ----
plot_data <- q21_tab %>%
  mutate(Response = str_wrap(as.character(q21_std), 44))

# ---- 3) Horizontal bar chart ----
p_q21 <- ggplot(plot_data, aes(
  y = fct_reorder(Response, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels (Count + Percent)
  geom_text(
    aes(label = Label),
    hjust = -0.05,
    fontface = "bold",
    size = 3.6,
    color = "black"
  ) +

  # X-axis
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles and caption
  labs(
    title    = "Q21: Confidence in Knowledge of AI Tools",
    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Confidence Level",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Publication-style theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1, face = "bold"),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 100, 35, 15)   # prevents right label clipping
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q21_confidence_in_AI_tools_horizontal_pub.png",
  plot = p_q21,
  dpi = 700,
  width = 9.0,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot & summary table ----
print(p_q21)

knitr::kable(
  q21_tab %>% rename(Response = q21_std),
  caption = paste0("Q21: Confidence in Knowledge of AI Tools (n = ", n_valid, ")")
)
```


## Q22: What types of training or support would you need to effectively integrate AI tools into your teaching?

```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

opt_map <- c(
  "1" = "More AI-related training or professional development",
  "2" = "Access to advanced AI platforms for classroom use",
  "3" = "Technical support for integrating AI into the curriculum",
  "4" = "Institutional support for purchasing AI tools",
  "5" = "Peer collaboration and sharing best practices"
)
canon_levels <- unname(opt_map)

# ---- Detect export style ----
nm <- names(faculty_analysis_data)
q22_single <- nm[tolower(nm) == "q22"]
q22_wide   <- nm[str_detect(tolower(nm), "^q22(\\.|_|\\[|:)?\\d+")]
q22_wide   <- setdiff(q22_wide, nm[str_detect(tolower(nm), "text|comment")])

faculty_analysis_data <- mutate(faculty_analysis_data, .row = row_number())

if (length(q22_wide) >= 1) {
  # Checkbox export: Q22_1 ... Q22_5
  long <- faculty_analysis_data %>%
    select(.row, all_of(q22_wide)) %>%
    pivot_longer(-.row, names_to = "col", values_to = "val") %>%
    mutate(
      idx = str_extract(col, "\\d+$"),
      option = unname(opt_map[idx]),
      val_chr = tolower(str_squish(as.character(val))),
      selected = case_when(
        val_chr %in% c("1","true","t","yes","y","selected","checked") ~ TRUE,
        suppressWarnings(!is.na(as.numeric(val_chr)) & as.numeric(val_chr) > 0) ~ TRUE,
        TRUE ~ FALSE
      )
    ) %>%
    filter(!is.na(option), selected) %>%
    select(.row, option)

} else if (length(q22_single) == 1) {
  # Single delimited column export: "A; B; C"
  long <- faculty_analysis_data %>%
    select(.row, q22 = all_of(q22_single)) %>%
    filter(!is.na(q22), q22 != "") %>%
    mutate(q22 = str_split(q22, "[,;]")) %>%
    unnest(q22) %>%
    mutate(q22 = str_squish(q22),
           option = case_when(
             str_detect(q22, regex("training|professional", TRUE)) ~ opt_map["1"],
             str_detect(q22, regex("platform|access|classroom", TRUE)) ~ opt_map["2"],
             str_detect(q22, regex("technical support|integrat|curriculum", TRUE)) ~ opt_map["3"],
             str_detect(q22, regex("institutional|purchas|tools", TRUE)) ~ opt_map["4"],
             str_detect(q22, regex("peer|collaboration|best practice", TRUE)) ~ opt_map["5"],
             TRUE ~ NA_character_
           )) %>%
    filter(!is.na(option)) %>%
    select(.row, option)
} else {
  stop("Q22 columns not found.")
}

# ---- Summary table (instrument order) ----
denom <- dplyr::n_distinct(long$.row)  # respondents who selected at least one Q22 item

q22_tab <- long %>%
  count(option, name = "Count") %>%
  complete(option = canon_levels, fill = list(Count = 0)) %>%
  mutate(
    option  = factor(option, levels = canon_levels),
    Percent = round(100 * Count / pmax(denom, 1), 1)
  ) %>%
  arrange(option)

knitr::kable(q22_tab, caption = "Q22: Training/Support Needs — Summary")

# ---- Plots ----
plot_data <- q22_tab %>%
  mutate(option_w = factor(str_wrap(as.character(option), 40),
                           levels = str_wrap(canon_levels, 40)))


# ---- 1) Sort and wrap long labels ----
plot_data <- q22_tab %>%
  mutate(
    option_w = fct_reorder(str_wrap(as.character(option), 40), Count)
  )

# ---- 1) Compute counts, percents, and labels ----
# ---- 1) Define denominator = number of unique respondents ----
denom <- 13  # or replace with n_distinct(faculty_analysis_data$.row) if dynamically computed

# ---- 2) Compute counts, percents, and labels ----
q22_faculty_tab <- plot_data %>%
  mutate(
    Percent = Count / denom,  # percentage of respondents
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Count))

# ---- 3) Horizontal bar chart (Counts + Percents) ----
p_q22_faculty_counts <- ggplot(q22_faculty_tab, aes(
  y = fct_reorder(option_w, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels (Count + Percent)
  geom_text(
    aes(label = Label),
    hjust = -0.05,
    fontface = "bold",
    size = 3.6,
    color = "black"
  ) +

  # X-axis formatting (percent of faculty)
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(q22_faculty_tab$Percent) + 0.15),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles & caption
  labs(
    title    = "Faculty Training and Support Needs for AI Integration",
    subtitle = "Multiple selections allowed; percentages based on faculty respondents (n = 13)",
    x        = "Percentage of Faculty",
    y        = "Training or Support Area",
    caption  = "Percentages represent proportion of faculty (not responses)."
  ) +

  # Publication theme (consistent with Q18–Q20)
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1, face = "bold"),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5, color = "gray30"),
    plot.caption = element_text(size = 9, hjust = 0, color = "gray40"),
    plot.margin = margin(10, 100, 35, 15)   # extra right margin for long labels
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q22_faculty_training_needs.png",
  plot = p_q22_faculty_counts,
  dpi = 700,
  width = 9.0,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot & summary table ----
print(p_q22_faculty_counts)

knitr::kable(
  q22_faculty_tab %>%
    rename(`Training/Support Area` = option_w) %>%
    select(`Training/Support Area`, Count, Percent),
  caption = "Q22: Faculty Training and Support Needs (Percent of Faculty, n = 13)"
)
```


## Q23: Do you believe your institution provides sufficient guidance and resources for using AI in teaching?


```{r, warning=FALSE, message=FALSE}

faculty_analysis_data <- if (exists("faculty_data_consented")) faculty_data_consented else faculty_data

q23_col <- dplyr::case_when(
  "Q23" %in% names(faculty_analysis_data) ~ "Q23",
  "q23" %in% names(faculty_analysis_data) ~ "q23",
  TRUE ~ NA_character_
)
stopifnot(!is.na(q23_col))

canon_levels <- c("Yes", "No", "Somewhat")

faculty_analysis_data <- faculty_analysis_data %>%
  mutate(
    q23_raw = .data[[q23_col]],
    q23_std = case_when(
      is.numeric(q23_raw) & q23_raw == 1 ~ "Yes",
      is.numeric(q23_raw) & q23_raw == 2 ~ "No",
      is.numeric(q23_raw) & q23_raw == 3 ~ "Somewhat",
      str_detect(as.character(q23_raw), regex("^yes$", ignore_case = TRUE)) ~ "Yes",
      str_detect(as.character(q23_raw), regex("^no$",  ignore_case = TRUE)) ~ "No",
      str_detect(as.character(q23_raw), regex("somewhat", ignore_case = TRUE)) ~ "Somewhat",
      TRUE ~ NA_character_
    ) |> factor(levels = canon_levels)
  )

# ---- Summary table ----
q23_ans <- faculty_analysis_data %>% filter(!is.na(q23_std))
denom   <- nrow(q23_ans)

q23_tab <- q23_ans %>%
  count(q23_std, name = "Count") %>%
  tidyr::complete(q23_std = factor(canon_levels, levels = canon_levels),
                  fill = list(Count = 0)) %>%
  mutate(Percent = round(100 * Count / pmax(denom, 1), 1),
         q23_std = factor(q23_std, levels = canon_levels))

knitr::kable(q23_tab,
             col.names = c("Response", "Count", "Percent"),
             caption   = "Q23: Does your institution provide sufficient AI guidance/resources? — Summary")

q23_tab <-  q23_ans  %>%
  filter(!is.na(q23_std) & q23_std != "") %>%
  count(q23_std, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Label   = paste0(Count, " (", round(Percent * 100, 1), "%)")
  ) %>%
  arrange(desc(Percent))

n_valid <- sum(q23_tab$Count)

# ---- 2) Prepare plot data with wrapped labels ----
plot_data <- q23_tab %>%
  mutate(Response = str_wrap(as.character(q23_std), 44))

# ---- 3) Horizontal bar chart (publication style) ----
p_q23 <- ggplot(plot_data, aes(
  y = fct_reorder(Response, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels (counts + percents)
  geom_text(
    aes(label = Label),
    hjust = -0.05,
    fontface = "bold",
    size = 3.6,
    color = "black"
  ) +

  # X-axis (percent scale)
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_data$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.05))
  ) +

  # Titles & caption
  labs(
    title    = "Q23: Do you believe your institution provides \nsufficient guidance and resources for using AI in teaching?",
    subtitle = "Single response per faculty; percentages add to 100%",
    x        = "Percentage of Faculty",
    y        = "Response",
    caption  = paste0("n = ", n_valid, " valid faculty responses.")
  ) +

  # Publication-style theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y  = element_text(size = 10.5, lineheight = 1.1),
    axis.text.x  = element_text(size = 10.5),
    plot.title   = element_text(face = "bold", size = 14),
    plot.subtitle= element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin  = margin(10, 100, 35, 15)   # extra right margin for labels
  )

# ---- 4) Save high-resolution figure ----
if (!dir.exists("Faculty_figures")) dir.create("Faculty_figures", recursive = TRUE)

ggsave(
  filename = "Faculty_figures/Q23_Institutional_Guidance.png",
  plot = p_q23,
  dpi = 700,
  width = 9.0,
  height = 5.5,
  units = "in"
)

# ---- 5) Display plot & summary table ----
print(p_q23)
knitr::kable(
  q23_tab %>% rename(Response = q23_std),
  caption = paste0("Q23: Institutional Guidance/Resources (n = ", n_valid, ")")
)
```


