---
output:
  html_document: default
  pdf_document: default
---

# Reproducible analysis of the student AI for data science survey
```{r setup-student-analysis, warning=FALSE, message=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
  dplyr, janitor, ggplot2, rlang, ggmosaic, knitr, forcats, tidyr,
  stringr, psych, readxl, scales, kableExtra, grid, tibble, lavaan
)

# Create output directory for saved figures
if (!dir.exists("Student_figures")) dir.create("Student_figures", recursive = TRUE)

find_student_file <- function() {
  candidates <- c(
    "AI4DS_Student_Survey_October_5_ 2025.xlsx",
    file.path("data", "AI4DS_Student_Survey_October_5_ 2025.xlsx")
  )
  found <- candidates[file.exists(candidates)]
  if (length(found) == 0) {
    stop(
      "Student survey Excel file not found in the working directory. ",
      "Place the workbook in the repo root or update `find_student_file()`."
    )
  }
  found[[1]]
}

load_student_data <- function(path = find_student_file(), min_duration = 120, min_progress = 25) {
  raw_preview <- readxl::read_excel(path)
  short <- readxl::read_excel(path, range = "A1:BN1", col_names = FALSE) |> unlist() |> as.character()
  long  <- readxl::read_excel(path, range = "A2:BN2", col_names = FALSE) |> unlist() |> as.character()
  n_cols <- ncol(raw_preview)
  short <- short[seq_len(n_cols)]
  long  <- long[seq_len(n_cols)]
  student_df <- readxl::read_excel(path, skip = 2, col_names = short)
  student_header_labels <- setNames(long, short)
  filtered_df <- student_df |>
    dplyr::filter(`Duration (in seconds)` >= min_duration, Progress >= min_progress)
  list(student_header_preview = raw_preview, student_data_raw = student_df, student_data_filtered = filtered_df, student_header_labels = student_header_labels)
}
```

```{r import-student-data, warning=FALSE, message=FALSE}
student_data_path <- find_student_file()
student_import <- load_student_data(student_data_path)
student_header_preview <- student_import$student_header_preview
student_data_raw <- student_import$student_data_raw
student_data_filtered <- student_import$student_data_filtered
student_header_labels <- student_import$student_header_labels

cat("Remaining sample size after quality filters:", nrow(student_data_filtered), "students\n")
dim(student_data_filtered)
```


## set up for the colors
```{r , warning=FALSE, message=FALSE}
## use this palette for all next figures
survey_palette_red_teal <- c("#117A65","#48C9B0","#F1C40F","#E67E22","#8B0000" )


```

## Saving custom theme for future use
```{r}
# --- Define once at the top of your script or in a separate R file ---
theme_publication <- function(base_size = 10) {
  theme_classic(base_size = base_size) +
    theme(
      plot.title      = element_text(face = "bold", size = 11, hjust = 0.5),
      plot.subtitle   = element_text(size = 10, hjust = 0.5, color = "gray25"),
      axis.text.x     = element_text(face = "bold", size = 10),
      axis.text.y     = element_text(face = "bold", size = 10, lineheight = 1.1),
      axis.title.x    = element_text(face = "bold", size = 10),
      axis.title.y    = element_text(face = "bold", size = 10),
      panel.grid.major.x = element_line(color = "gray90"),
      panel.grid.minor = element_blank(),
      legend.title    = element_text(face = "bold"),
      legend.text     = element_text(size = 8),
      legend.position = "bottom",
      legend.key.size = grid::unit(0.35, "cm"),
      legend.margin   = margin(t = 2, b = 5),
      legend.box.margin = margin(5, 5, 10, 5),
      legend.box = "vertical",
      plot.caption    = element_text(size = 9, hjust = 0, color = "gray30"),
      plot.margin     = margin(10, 15, 45, 15)
    )
}

###
load_publication_style <- function() {
  # Color palette
  assign("familiarity_palette", c(
    "Not at All Familiar" = "#B2182B",
    "Not Very Familiar"   = "#E67E22",
    "Neutral"             = "#F1C40F",
    "Somewhat Familiar"   = "#48C9B0",
    "Very Familiar"       = "#117A65"
  ), envir = .GlobalEnv)

  # Theme
  theme_set(theme_publication())
}

load_publication_style()

```

## Demographics sex table Q3-6
```{r, warning=FALSE, message=FALSE}
## Demographics table Q3-6
# ---- 1) Q3 Summary: Student Sex ----
q3_summary <- student_data_filtered %>%
  filter(!is.na(Q3)) %>%
  tabyl(Q3) %>%
  adorn_totals("row") %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 3)


# ---- 1) Q3 Summary: Student Sex ----


# total valid responses
n_valid_q3 <- student_data_filtered %>% filter(!is.na(Q3)) %>% nrow()

q3_summary <- student_data_filtered %>%
  filter(!is.na(Q3)) %>%
  tabyl(Q3) %>%                 # gives Q3 + n
  adorn_totals("row") %>%       # adds Total row
  mutate(
    percent = n / n_valid_q3,   # keep counts in `n`, compute percent separately
    percent = percent(percent, accuracy = 0.1)  # format, e.g., "61.2%"
  )

q3_summary

# Display as a formatted table
kable(q3_summary, caption = "Q3: Student Sex - Counts and Percentages") %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 13)
##

# Optional bar plot
ggplot(student_data_filtered %>% filter(!is.na(Q3)), aes(x = Q3)) +
  geom_bar(fill = "#48C9B0", color = "gray30", width = 0.7) +
   coord_flip()+
  geom_text(stat = "count", aes(label = ..count..), vjust = 1, fontface = "bold") +
  labs(
    title = "Student Sex",
    x = NULL, y = "Count"
  ) +
  theme_publication() +
  theme(
    axis.text.x = element_text(face = "bold"),
    plot.margin = margin(10, 20, 25, 15)
  )


```

## major
```{r, warning=FALSE, message=FALSE}
# ---- 2) Q4 Summary (e.g., Major, Class Level, or Other Demographic) ----
q4_summary <- student_data_filtered %>%
  filter(!is.na(Q4)) %>%
  tabyl(Q4) %>%
  adorn_totals("row") %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 3)

# ---- 1) Clean and consolidate major names ----
cleaned_majors <- student_data_filtered %>%
  filter(!is.na(Q4)) %>%
  mutate(
    major_clean = str_to_lower(Q4),  # make lowercase
    major_clean = str_squish(major_clean),  # remove double spaces

    # Consolidate common variations
    major_group = case_when(
      str_detect(major_clean, "stat") |
      str_detect(major_clean, "data") |
      str_detect(major_clean, "analytic") ~ "Statistics & Data Science",

      str_detect(major_clean, "computer") |
      str_detect(major_clean, "cs") |
      str_detect(major_clean, "software") ~ "Computer Science / IT",

      str_detect(major_clean, "engineer") ~ "Engineering (All Types)",

      str_detect(major_clean, "math") ~ "Mathematics",

      str_detect(major_clean, "bio") |
      str_detect(major_clean, "life") |
      str_detect(major_clean, "chem") |
      str_detect(major_clean, "enviro") ~ "Biological / Physical Sciences",

      str_detect(major_clean, "business") |
      str_detect(major_clean, "finance") |
      str_detect(major_clean, "account") |
      str_detect(major_clean, "econ") ~ "Business / Economics",

      str_detect(major_clean, "psych") ~ "Psychology",
      str_detect(major_clean, "edu") ~ "Education",
      str_detect(major_clean, "nurs") | str_detect(major_clean, "health") ~ "Health Sciences",

      TRUE ~ "Other / Undeclared"
    )
  )

# ---- 2) Summarize counts and percentages ----
major_summary <- cleaned_majors %>%
  count(major_group, name = "Count") %>%
  mutate(
    Percent = Count / sum(Count),
    Percent_Label = paste0(round(Percent * 100, 1), "%")
  ) %>%
  arrange(desc(Count))

# ---- 3) Clean, publication-ready table ----
kable(
  major_summary,
  caption = "Q4: Student Majors (Consolidated Categories)",
  col.names = c("Major Category", "Count", "Percent", "Percent Label"),
  digits = 1,
  align = "lccr"
) %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 12)

# ---- 4) Optional: Horizontal bar chart ----
ggplot(major_summary, aes(y = fct_reorder(major_group, Count), x = Count)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +
  geom_text(aes(label = paste0(Count, " (", round(Percent * 100, 1), "%)")),
            hjust = -0.1, fontface = "bold", size = 3.6, color = "black") +
  labs(
    title = "Q4: Student Majors (Consolidated Categories)",
    x = "Number of Students",
    y = "Major Category"
  ) +
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 35, 15)
  )

# ---- 5) (Optional) Save the figure ----
ggsave(
  filename = "Student_figures/Q4_student_majors_consolidated.png",
  dpi = 600,
  width = 9.5,
  height = 5.5,
  units = "in"
)

```

## Demographics table Q3-6
```{r, warning=FALSE, message=FALSE}
## Demographics table Q3-6


student_data_filtered <- student_data_filtered %>%
  mutate(
    Q5 = factor(Q5, levels = c("Freshman", "Sophomore", "Junior", "Senior", "Graduate Student")))


# 4) Q5 summary now includes former NAs as "Graduate"
student_data_filtered %>% tabyl(Q5)

# 5) Q5 bar chart (no NA filter needed; includes "Graduate")
student_data_filtered %>%
  count(Q5) %>%
  ggplot(aes(x = Q5, y = n)) +
  geom_bar(stat = "identity", fill = "#48C9B0") +
  ggtitle("What is your classification?") +
  labs(x = "Classification", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Q6
```{R}
# ---- 1) Define ordered factor levels ----
student_data_filtered$Q6 <- factor(
  student_data_filtered$Q6,
  levels = c("Never", "Rarely", "Sometimes", "Often", "Very Often")
)

# ---- 2) Compute frequencies and percentages ----
plot_df_q6 <- student_data_filtered %>%
  filter(!is.na(Q6)) %>%
  count(Q6) %>%
  mutate(Percent = n / sum(n))

# ---- 3) Horizontal bar chart with publication theme ----
p_q6 <- ggplot(plot_df_q6, aes(y = fct_reorder(Q6, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q6$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and labels
  labs(
    title = "Q6: Frequency of Using AI Technologies in Data Science Work",
    subtitle = "Includes tools such as ChatGPT, Copilot, and Google Colab",
    x = "Percentage of Respondents",
    y = "Frequency",
    caption = "Note: Percentages calculated from valid (non-missing) responses."
  ) +

  # Apply unified publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# ---- 4) Save as high-resolution PNG ----
ggsave(
  filename = "Student_figures/Q6_ai_usage_frequency_horizontal.png",
  plot = p_q6,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q6)

```

## Q7 ## percnetage of all mentions in responses
```{r}
## student_data_filtered %>% tabyl(Q7)
## needs wrangling cuz we allowed multiple responses
# Step 1: Split comma-separated responses into rows
long_Q7 <- student_data_filtered %>%
  select(Q7) %>%
  filter(!is.na(Q7)) %>%  ## remove NAs
  mutate(Q7 = str_split(Q7, ",")) %>%
  unnest(Q7)

# Step 2: Clean up whitespace
long_Q7 <- long_Q7 %>%
  mutate(Q7 = str_trim(Q7))

# Step 3: Count frequencies
Q7_summary <- long_Q7 %>%
  count(Q7, sort = TRUE) %>%
  rename(Response = Q7, Frequency = n)
# View summary
print(Q7_summary)

## Frequency of use case as percent of 82 responses
# Step 3: Count frequencies and compute relative frequencies by total respondents (82)
Q7_summary2 <- long_Q7 %>%
  count(Q7, sort = TRUE) %>%
  rename(Response = Q7, Frequency = n) %>%
  mutate(Relative_Freq = round(Frequency / 120*100, digits=1))

# View summary
print(Q7_summary2)


```

## Q7 Percents based on valid responses
```{r}
# ---- 1) Split and clean comma-separated responses ----
long_Q7 <- student_data_filtered %>%
  select(Q7) %>%
  filter(!is.na(Q7)) %>%
  mutate(Q7 = str_split(Q7, ",")) %>%
  unnest(Q7) %>%
  mutate(Q7 = str_trim(Q7))

# ---- 2) Count responses and compute percent based on valid respondents ----
n_valid <- nrow(student_data_filtered %>% filter(!is.na(Q7)))

Q7_summary <- long_Q7 %>%
  count(Q7, sort = TRUE) %>%
  rename(Response = Q7, Count = n) %>%
  mutate(
    Percent = Count / n_valid,
    Label = paste0(Count, " (", round(Percent * 100, 1), "%)")
  )

# ---- 3) Horizontal bar chart with counts + percentages ----
p_q7 <- ggplot(Q7_summary, aes(
  y = fct_reorder(Response, Percent),
  x = Percent
)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline labels showing count and percent
  geom_text(
    aes(label = Label),
    hjust = -0.1, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(Q7_summary$Percent) + 0.1),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and labels
  labs(
    title = "Q7: Common Use Cases for AI Tools in Data Science Tasks",
    subtitle = "Multiple selections allowed; percentages based on valid respondents",
    x = "Percentage of Respondents Selecting Each Use Case",
    y = "Use Case",
    caption = paste0("n = ", n_valid,
                     " respondents (percentages may exceed 100% because multiple choices were allowed).")
  ) +

  # Minimal publication-style theme
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    plot.caption = element_text(size = 9),
    plot.margin = margin(10, 25, 25, 15)
  )

# Display chart
p_q7

```

## Q8: Are you interested in pursuing a degree in data science in the future?
```{r}
# ---- 1) Define ordered factor levels ----
student_data_filtered$Q8 <- factor(
  student_data_filtered$Q8,
  levels = c("Yes", "No", "Not sure")
)

# ---- 2) Compute frequencies and percentages ----
plot_df_q8 <- student_data_filtered %>%
  filter(!is.na(Q8)) %>%
  count(Q8) %>%
  mutate(Percent = n / sum(n))

# ---- 3) Horizontal bar chart (publication theme) ----
p_q8 <- ggplot(plot_df_q8, aes(y = fct_reorder(Q8, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q8$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and labels
  labs(
    title = "Q8: Interest in Pursuing a Degree in Data Science",
    subtitle = "Responses show future interest in formal data science education",
    x = "Percentage of Respondents",
    y = "Response",
    caption = "Note: Percentages calculated from valid (non-missing) responses."
  ) +

  # Unified publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# ---- 4) Save as high-resolution PNG ----
ggsave(
  filename = "Student_figures/Q8_interest_in_ds_degree_horizontal.png",
  plot = p_q8,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q8)


```

## Q9: Are you interested in getting a job in data science in the future?

```{r}
# ---- 1) Define response order ----
student_data_filtered$Q9 <- factor(
  student_data_filtered$Q9,
  levels = c("Yes", "No", "Not sure")
)

# ---- 2) Compute frequencies and percentages ----
plot_df_q9 <- student_data_filtered %>%
  filter(!is.na(Q9)) %>%
  count(Q9) %>%
  mutate(Percent = n / sum(n))

# ---- 3) Horizontal bar chart with publication theme ----
p_q9 <- ggplot(plot_df_q9, aes(y = fct_reorder(Q9, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q9$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and captions
  labs(
    title = "Q9: Interest in Future Careers in Data Science",
    subtitle = "Responses show majority interest in pursuing data science roles",
    x = "Percentage of Respondents",
    y = "Response",
    caption = "Note: Percentages based on valid (non-missing) responses."
  ) +

  # Unified publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# ---- 4) Save as PNG ----
ggsave(
  filename = "Student_figures/Q9_interest_in_ds_career_horizontal.png",
  plot = p_q9,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q9)

```

## Q10: How familiar are you with the following AI technologies in data science?

```{r}
# Shared Levels and Palette
levels_q10 <- c("Not at All Familiar", "Not Very Familiar", "Neutral", "Somewhat Familiar", "Very Familiar")

familiarity_palette <- c(
  "Not at All Familiar" = "#B2182B",
  "Not Very Familiar"   = "#E67E22",
  "Neutral"             = "#F1C40F",
  "Somewhat Familiar"   = "#48C9B0",
  "Very Familiar"       = "#117A65"
)
#Labels and Levels
tech_map <- c(
  Q10_1 = "ChatGPT",
  Q10_2 = "Gemini",
  Q10_3 = "Google Colab",
  Q10_4 = "AI-powered\ndata visualization tools",
  Q10_5 = "AI-driven\nanalytics platforms"
)

# Data Preparation

df_long_q10 <- student_data_filtered %>%
  pivot_longer(starts_with("Q10_"), names_to = "Item", values_to = "Response") %>%
  mutate(
    Technology = recode(Item, !!!tech_map, .default = NA_character_),
    Response   = factor(Response, levels = levels_q10)
  ) %>%
  filter(!is.na(Response), !is.na(Technology))

plot_df_q10 <- df_long_q10 %>%
  count(Technology, Response, .drop = FALSE) %>%
  group_by(Technology) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  group_by(Technology) %>%
  mutate(pct_familiar = sum(pct[Response %in% c("Somewhat Familiar", "Very Familiar")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Technology = fct_reorder(Technology, pct_familiar, .desc = TRUE))


p_q10 <- ggplot(plot_df_q10, aes(y = Technology, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = familiarity_palette, guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Q10: Student Familiarity with AI Technologies in Data Science",
    subtitle = "Responses grouped by self-reported familiarity level",
    x = "Percentage of Responses",
    y = "AI Technology",
    fill = "Response",
    caption = "Note: Percentages represent the proportion of responses within each technology category."
  ) +
  theme_publication()


print(p_q10)
ggsave(
  filename = "Student_figures/Q10_familiarity_horizontal_fxd.png",
  plot = p_q10,
  dpi = 700,
  width = 9.5,
  height = 6.5,  # slightly taller for legend space
  units = "in"
)


```

## Q11: Which AI tools have you used in your data science tasks? (Select all that apply) - Selected Choice

```{R}
#Define the exact expected options (based on your survey config)
Q11_options <- c(
  "ChatGPT",
  "Grammarly",
  "Claude and Copilot",
  "AI-powered visualization tools",
  "Copilot",
  "Gemini",
  "Canva or AI assisted presentation tools",
  "Predictive analytics tools (e.g.DataRobot)",
  "None",
  "Others please specify"
)

#Extract responses and clean
long_Q11 <- student_data_filtered %>%
  select(Q11) %>%
  filter(!is.na(Q11))

#Create a binary indicator for each known option
for (opt in Q11_options) {
  long_Q11[[opt]] <- str_detect(long_Q11$Q11, fixed(opt))
}

#Convert wide to long and count
Q11_summary <- long_Q11 %>%
  pivot_longer(cols = all_of(Q11_options), names_to =   "Response", values_to = "Selected") %>%
  filter(Selected) %>%
  count(Response, sort = TRUE)

print(Q11_summary)

## With Percents
Q11_summary2 <- long_Q11 %>%
  count(Q11, sort = TRUE) %>%
  rename(Response = Q11, Frequency = n) %>%
  mutate(Relative_Freq = round(Frequency / 120*100, digits=1))

# summary
print(Q11_summary2)
```

```{r}
# Define expected options
Q11_options <- c(
  "ChatGPT",
  "Grammarly",
  "Claude and Copilot",
  "AI-powered visualization tools",
  "Copilot",
  "Gemini",
  "Canva or AI assisted presentation tools",
  "Predictive analytics tools (e.g.DataRobot)",
  "None",
  "Others please specify"
)

#Extract and clean responses
long_Q11 <- student_data_filtered %>%
  select(Q11) %>%
  filter(!is.na(Q11))

#Create binary indicators for each tool
for (opt in Q11_options) {
  long_Q11[[opt]] <- str_detect(long_Q11$Q11, fixed(opt))
}

# Convert wide → long and summarize
Q11_summary <- long_Q11 %>%
  pivot_longer(
    cols = all_of(Q11_options),
    names_to = "Tool",
    values_to = "Selected"
  ) %>%
  filter(Selected) %>%
  count(Tool, name = "Count") %>%
  mutate(Percent = Count / 120)

# Polished horizontal bar chart
p_q11 <- ggplot(Q11_summary, aes(y = fct_reorder(Tool, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis settings
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(Q11_summary$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and captions
  labs(
    title = "Q11: AI Tools Used in Data Science Tasks",
    subtitle = "Multiple selections allowed; shows percentage of respondents selecting each tool",
    x = "Percentage of Respondents",
    y = "AI Tool",
    caption = "Note: Percentages calculated from valid responses."
  ) +

  # Apply your publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# Save as PNG
ggsave(
  filename = "Student_figures/Q11_ai_tools_used_horizontal.png",
  plot = p_q11,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q11)

```

## Q12_1: To what extent do you agree with the following statements regarding AI tools in data science?

```{r}
# Map Q12_
statement_map <- c(
  Q12_1 = "AI tools enhance \naccuracy",
  Q12_2 = "AI tools automate \nprocessing",
  Q12_3 = "AI tools provide \nnew insights",
  Q12_4 = "AI tools lack \nemotional intelligence"
)

# Long format + recode
df_long_q12 <- student_data_filtered %>%
  pivot_longer(
    cols = starts_with("Q12_"),
    names_to = "Item",
    values_to = "Response"
  ) %>%
  mutate(
    Statement = dplyr::recode(Item, !!!statement_map, .default = NA_character_),
    Response = factor(
      Response,
      levels = c("Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree")
    )
  ) %>%
  filter(!is.na(Response), !is.na(Statement))

# Compute proportions per statement
plot_df_q12 <- df_long_q12 %>%
  count(Statement, Response, .drop = FALSE) %>%
  group_by(Statement) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

#  Order statements by % favorable (Agree + Strongly Agree)
plot_df_q12 <- plot_df_q12 %>%
  group_by(Statement) %>%
  mutate(pct_fav = sum(pct[Response %in% c("Agree", "Strongly Agree")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Statement = fct_reorder(Statement, pct_fav, .desc = TRUE))

#Custom palette
pal_q12 <- c(
  "Strongly Disagree" = "#B2182B",
  "Disagree"          = "#E67E22",
  "Neutral"           = "#F1C40F",
  "Agree"             = "#48C9B0",
  "Strongly Agree"    = "#117A65"
)

# Horizontal stacked proportion bar chart
p_q12 <- ggplot(plot_df_q12, aes(y = Statement, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +

  # Optional percentage labels
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +

  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = pal_q12, guide = guide_legend(reverse = TRUE)) +

  labs(
    title = "Q12: Attitudes Toward AI Tools in Data Science",
    subtitle = "Responses grouped by agreement level",
    x = "Percentage of Responses",
    y = "Statement",
    fill = "Response",
    caption = "Note: Percentages represent the proportion of responses within each statement."
  ) +

  # saved publication theme
  theme_publication()

# High-Resolution Figure
ggsave(
  filename = "Student_figures/Q12_attitudes_horizontal_pub.pdf",
  plot = p_q12,
  dpi = 600,
  width = 9,
  height = 6,
  units = "in"
)

print(p_q12)


```

## Q13: “How often do you use AI tools for the following data science tasks? (Never, Rarely, Sometimes, Often, Always)”

```{r}

# Map Q13_*
task_map <- c(
  Q13_1 = "Exploratory data analysis",
  Q13_2 = "Predictive modeling",
  Q13_3 = "Data visualization",
  Q13_4 = "Automating data cleaning tasks",
  Q13_5 = "Presenting findings and reports"
)

# Define levels + consistent palette
levels_q13 <- c("Never", "Rarely", "Sometimes", "Often", "Always")
pal_q13 <- c(
  "Never"     = "#B2182B",  # red
  "Rarely"    = "#E67E22",  # orange
  "Sometimes" = "#F1C40F",  # gold
  "Often"     = "#48C9B0",  # turquoise
  "Always"    = "#117A65"   # teal green
)

#Reshape and clean
df_long_q13 <- student_data_filtered %>%
  pivot_longer(
    cols = starts_with("Q13_"),
    names_to = "Item",
    values_to = "Response"
  ) %>%
  mutate(
    Task     = recode(Item, !!!task_map, .default = NA_character_),
    Response = factor(Response, levels = levels_q13)
  ) %>%
  filter(!is.na(Response), !is.na(Task))

# Calculate proportions
plot_df_q13 <- df_long_q13 %>%
  count(Task, Response, .drop = FALSE) %>%
  group_by(Task) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

#Order tasks by % high use
plot_df_q13 <- plot_df_q13 %>%
  group_by(Task) %>%
  mutate(pct_high = sum(pct[Response %in% c("Often", "Always")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Task = fct_reorder(Task, pct_high, .desc = TRUE))

# Horizontal stacked bar chart
p_q13 <- ggplot(plot_df_q13, aes(y = Task, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +

  # Optional: add centered % labels inside bars
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +

  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  scale_fill_manual(values = pal_q13, guide = guide_legend(reverse = TRUE)) +

  labs(
    title = "Q13: How often do you use AI tools for data science tasks?",
    subtitle = "Responses grouped by frequency of use",
    x = "Percentage of Responses",
    y = "Task",
    fill = "Response",
    caption = "Note: Percentages represent the proportion of responses within each task."
  ) +

  #publication theme
  theme_publication()

# high-resolution figure
ggsave(
  filename = "Student_figures/Q13_utilization_horizontal_pub.pdf",
  plot = p_q13,
  dpi = 600,
  width = 9,
  height = 6,
  units = "in"
)

print(p_q13)


```

## Q14: How confident are you in explaining AI-generated outputs for data analysis tasks to peers or supervisors?

```{r}
student_data_filtered$Q14 <- factor(student_data_filtered$Q14,
  levels = c("Very Confident", "Confident", "Neutral",
             "Not Very Confident", "Not at All Confident"))

# Frequency table
student_data_filtered %>%
  filter(!is.na(Q14)) %>%
  tabyl(Q14)


```

```{r}
# Factor order and consistent color palette
conf_levels <- c(
  "Not at All Confident",
  "Not Very Confident",
  "Neutral",
  "Confident",
  "Very Confident"
)

conf_colors <- c(
  "Not at All Confident" = "#B2182B",
  "Not Very Confident"   = "#E67E22",
  "Neutral"              = "#F1C40F",
  "Confident"            = "#48C9B0",
  "Very Confident"       = "#117A65"
)


plot_df_q14 <- student_data_filtered %>%
  filter(!is.na(Q14)) %>%
  mutate(Response = factor(Q14, levels = conf_levels)) %>%
  count(Response, .drop = FALSE) %>%
  mutate(
    pct   = n / sum(n),
    label = scales::percent(pct, accuracy = 0.1)
  )


p_q14 <- ggplot(plot_df_q14,
                aes(y = "Confidence in \nExplaining AI Outputs", x = pct, fill = Response)) +
  geom_col(width = 0.6, color = "gray30") +

  # Percent labels inside segments (only for larger segments)
  geom_text(
    aes(label = ifelse(pct >= 0.03, label, "")),
    position = position_stack(vjust = 0.5),
    size = 3, fontface = "bold", color = "black"
  ) +

  # X-axis in percentage format
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +


  scale_fill_manual(values = conf_colors, guide = guide_legend(reverse = TRUE)) +

  # Titles, labels, caption
  labs(
    title = "Q14: Ability to Explain AI-Generated Outputs",
    subtitle = "Student self-reported confidence level",
    x = "Percentage of Responses",
    y = NULL,
    fill = "Response",
#    caption = "Note: Percentages represent the proportion of respondents selecting each confidence level."
  ) +

  theme_publication()

# high-resolution figure
ggsave(
  filename = "Student_figures/Q14_explainability_horizontal_pub.pdf",
  plot = p_q14,
  dpi = 600,
  width = 8.5,
  height = 4.5,
  units = "in"
)

print(p_q14)

```

## Q15: To what extent do you agree with the following statements regarding the future of AI in data science? (Strongly Disagree, Disagree, Neutral, Agree, Strongly Agree)

```{r}
#  map and consistent palette
statement_map_q15 <- c(
  Q15_1 = "AI tools essential \nfor future workflows",
  Q15_2 = "AI improves efficiency \nin large datasets",
  Q15_3 = "AI makes DS accessible \nto non-technical professionals",
  Q15_4 = "Overreliance on AI \nmay diminish skills"
)

levels_q15 <- c("Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree")
pal_q15 <- c(
  "Strongly Disagree" = "#B2182B",
  "Disagree"          = "#E67E22",
  "Neutral"           = "#F1C40F",
  "Agree"             = "#48C9B0",
  "Strongly Agree"    = "#117A65"
)

# Reshape and clean
df_long_q15 <- student_data_filtered %>%
  pivot_longer(
    cols = starts_with("Q15_"),
    names_to = "Item",
    values_to = "Response"
  ) %>%
  mutate(
    Statement = recode(Item, !!!statement_map_q15, .default = NA_character_),
    Response  = factor(Response, levels = levels_q15)
  ) %>%
  filter(!is.na(Response), !is.na(Statement))

# Calculate proportions
plot_df_q15 <- df_long_q15 %>%
  count(Statement, Response, .drop = FALSE) %>%
  group_by(Statement) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  group_by(Statement) %>%
  mutate(pct_agree = sum(pct[Response %in% c("Agree", "Strongly Agree")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Statement = fct_reorder(Statement, pct_agree, .desc = TRUE))


p_q15 <- ggplot(plot_df_q15, aes(y = Statement, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +


  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +

  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = pal_q15, guide = guide_legend(reverse = TRUE)) +

  labs(
    title = "Q15: Attitudes about the Future of AI in Data Science",
    subtitle = "Responses grouped by agreement level",
    x = "Percentage of Responses",
    y = "Statement",
    fill = "Response",
    caption = "Note: Percentages represent the proportion of responses within each statement."
  ) +

  # publication theme
  theme_publication()

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q15_attitudes_horizontal_pub.png",
  plot = p_q15,
  dpi = 600,
  width = 9,
  height = 6,
  units = "in"
)

print(p_q15)


```

## Q16: Do you think students should learn how to use AI tools for their data science careers?

```{r}
# Define ordered factor levels
student_data_filtered$Q16 <- factor(student_data_filtered$Q16,
  levels = c(
    "Yes, it's essential",
    "Yes, but not a priority",
    "Neutral",
    "No, it's not necessary",
    "No, it may even be harmful"
  )
)

# Compute frequencies and percentages
plot_df_q16 <- student_data_filtered %>%
  filter(!is.na(Q16)) %>%
  count(Q16) %>%
  mutate(Percent = n / sum(n))

# Horizontal bar chart (publication style)
p_q16 <- ggplot(plot_df_q16, aes(y = fct_reorder(Q16, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q16$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and captions
  labs(
    title = "Q16: Should Students Learn AI Tools for Their Data Science Careers?",
    subtitle = "Responses indicate strong support for AI tool proficiency in DS education",
    x = "Percentage of Respondents",
    y = "Response",
    caption = "Note: Percentages calculated from valid (non-missing) responses."
  ) +

  # Unified publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q16_ai_learning_importance_horizontal.png",
  plot = p_q16,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q16)

```

## Q17: To what extent do you agree with the following statement: I am willing to incorporate AI tools more into my data science projects in the future.

```{r}
#Define factor order
student_data_filtered$Q17 <- factor(student_data_filtered$Q17,
  levels = c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree")
)

# Compute frequency and percentages
plot_df_q17 <- student_data_filtered %>%
  filter(!is.na(Q17)) %>%
  count(Q17) %>%
  mutate(Percent = n / sum(n))

#Horizontal bar chart
p_q17 <- ggplot(plot_df_q17, aes(y = fct_reorder(Q17, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q17$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and labels
  labs(
    title = "Q17: Willingness to Incorporate AI Tools in Future Data Science Projects",
    subtitle = "Responses ordered by level of agreement",
    x = "Percentage of Respondents",
    y = "Response",
    caption = "Note: Percentages calculated from valid responses."
  ) +

  # Apply your shared publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q17_integration_horizontal.png",
  plot = p_q17,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q17)

```

## Q18: To what extent are you aware of the following limitations of using AI for data science tasks?  (Very Aware, Somewhat Aware, Neutral, Not Very Aware, Not at All Aware)

```{r}
# Define recode map
limitation_map <- c(
  Q18_1 = "Biased results",
  Q18_2 = "Factually incorrect outputs",
  Q18_3 = "Lack real-world \nexperience/context",
  Q18_4 = "Need validation \nby experts",
  Q18_5 = "Data privacy risks"
)

# factor levels + consistent palette
levels_q18 <- c(
  "Not at All Aware",
  "Not Very Aware",
  "Neutral",
  "Somewhat Aware",
  "Very Aware"
)

pal_q18 <- c(
  "Not at All Aware" = "#B2182B",
  "Not Very Aware"   = "#E67E22",
  "Neutral"          = "#F1C40F",
  "Somewhat Aware"   = "#48C9B0",
  "Very Aware"       = "#117A65"
)

# Reshape and clean
df_long_q18 <- student_data_filtered %>%
  pivot_longer(
    cols = starts_with("Q18_"),
    names_to = "Item",
    values_to = "Response"
  ) %>%
  mutate(
    Limitation = recode(Item, !!!limitation_map, .default = NA_character_),
    Response   = factor(Response, levels = levels_q18)
  ) %>%
  filter(!is.na(Response), !is.na(Limitation))

#Calculate proportions
plot_df_q18 <- df_long_q18 %>%
  count(Limitation, Response, .drop = FALSE) %>%
  group_by(Limitation) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

#Order by % high awareness
plot_df_q18 <- plot_df_q18 %>%
  group_by(Limitation) %>%
  mutate(pct_high = sum(pct[Response %in% c("Somewhat Aware","Very Aware")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Limitation = fct_reorder(Limitation, pct_high, .desc = TRUE))

# Horizontal stacked bar chart
p_q18 <- ggplot(plot_df_q18, aes(y = Limitation, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +

  # Optional percentage labels
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +

  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = pal_q18, guide = guide_legend(reverse = TRUE)) +

  labs(
    title = "Q18: Awareness of AI Limitations in Data Science",
    subtitle = "Responses grouped by awareness level",
    x = "Percentage of Responses",
    y = "AI Limitation",
    fill = "Response",
    caption = "Note: Percentages represent the proportion of responses within each limitation category."
  ) +

  theme_publication()

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q18_awareness_horizontal_pub.png",
  plot = p_q18,
  dpi = 600,
  width = 9,
  height = 6,
  units = "in"
)

print(p_q18)


```


## Q19: What are your primary concerns about using AI tools in data science?
```{r}
# Create new variables (binary indicators) for each concern
student_data_filtered <- student_data_filtered %>% mutate( concern_accuracy = str_detect(Q19, regex("accuracy", ignore_case = TRUE)), concern_transparency = str_detect(Q19, regex("transparency", ignore_case = TRUE)), concern_privacy = str_detect(Q19, regex("privacy", ignore_case = TRUE)), concern_ethics = str_detect(Q19, regex("ethic|bias", ignore_case = TRUE)), concern_skill_loss = str_detect(Q19, regex("skill", ignore_case = TRUE)), concern_understanding = str_detect(Q19, regex("understanding", ignore_case = TRUE)),
concern_none = str_detect(Q19, regex("don't|none", ignore_case = TRUE)) )
# Count totals for each concern
q19_summary <- tibble( Concern = c("Accuracy & Reliability", "Lack of Transparency", "Privacy & Security", "Ethical Concerns (Bias)", "Skill Degradation", "Difficulty Understanding Models", "Don't Know/None"),
                       Count = c( sum(student_data_filtered$concern_accuracy, na.rm = TRUE), sum(student_data_filtered$concern_transparency, na.rm = TRUE), sum(student_data_filtered$concern_privacy, na.rm = TRUE), sum(student_data_filtered$concern_ethics, na.rm = TRUE), sum(student_data_filtered$concern_skill_loss, na.rm = TRUE), sum(student_data_filtered$concern_understanding, na.rm = TRUE), sum(student_data_filtered$concern_none, na.rm = TRUE) ) )
# Add percentages
q19_summary <- q19_summary %>% mutate(Percent = round(Count / 120 * 100, 1))

# ---- 1) Binary indicators already created ----
# (as in your code)
(n19 <- sum(!is.na(student_data_filtered$Q19)))
# ---- 2) Summarize + add percentages ----
q19_summary <- tibble(
  Concern = c("Accuracy & Reliability",
              "Lack of Transparency",
              "Privacy & Security",
              "Ethical Concerns (Bias)",
              "Skill Degradation",
              "Difficulty Understanding Models",
              "Don't Know / None"),
  Count = c(
    sum(student_data_filtered$concern_accuracy, na.rm = TRUE),
    sum(student_data_filtered$concern_transparency, na.rm = TRUE),
    sum(student_data_filtered$concern_privacy, na.rm = TRUE),
    sum(student_data_filtered$concern_ethics, na.rm = TRUE),
    sum(student_data_filtered$concern_skill_loss, na.rm = TRUE),
    sum(student_data_filtered$concern_understanding, na.rm = TRUE),
    sum(student_data_filtered$concern_none, na.rm = TRUE)
  )
) %>%
  mutate(Percent = Count / n19)

# Polished horizontal bar chart
p_q19 <- ggplot(q19_summary, aes(y = fct_reorder(Concern, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline % labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, size = 3.6, fontface = "bold", color = "black"
  ) +

  # Axes formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(q19_summary$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and caption
  labs(
    title = "Q19: Primary Concerns About Using AI Tools in Data Science",
    subtitle = "Top issues mentioned by respondents",
    x = "Percentage of Respondents",
    y = "Concern",
    caption = "Note: Percentages calculated from valid (non-missing) responses"
  ) +

  # Unified publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

#Save as high-resolution PNG
ggsave(
  filename = "Student_figures/Q19_primary_concerns_horizontal.png",
  plot = p_q19,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q19)
```

## Q20: Have you ever encountered issues such as hallucinations, bias, or inaccuracy while using AI tools for data science tasks?


```{r}
# Define factor order
student_data_filtered$Q20 <- factor(student_data_filtered$Q20,
  levels = c(
    "Yes, frequently",
    "Yes, occasionally",
    "No, never",
    "I never used AI tools for data science"
  )
)

# Compute frequency and percentages
n20 = sum(!is.na(student_data_filtered$Q20))
plot_df_q20 <- student_data_filtered %>%
  filter(!is.na(Q20)) %>%
  count(Q20) %>%
  mutate(Percent = n / n20)


#  Horizontal bar chart
p_q20 <- ggplot(plot_df_q20, aes(y = fct_reorder(Q20, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percent labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q20$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles and captions
  labs(
    title = "Q20: Encountered Issues with AI Tools in Data Science Tasks",
    subtitle = "Frequency of reported challenges when using AI tools",
    x = "Percentage of Respondents",
    y = "Response",
    caption = "Note: Percentages calculated from valid (non-missing) responses."
  ) +


  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q20_ai_tool_issues_horizontal_pub.png",
  plot = p_q20,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q20)

```

## Q21: Do you think AI tools can significantly enhance the following aspects of your data science workflow?  (Strongly Disagree, Disagree, Neutral, Agree, Strongly Agree)


```{r}
#  Define recode map
aspect_map_q21 <- c(
  Q21_1 = "Automating repetitive \ntasks",
  Q21_2 = "Enhancing predictive \nmodeling accuracy",
  Q21_3 = "Visualizing \ncomplex datasets",
  Q21_4 = "Supporting faster \ndecision-making",
  Q21_5 = "Communicating to \nnon-technical stakeholders"
)

# ---- 2) Define factor levels + consistent palette ----
levels_q21 <- c("Strongly Disagree", "Disagree", "Neutral", "Agree", "Strongly Agree")
pal_q21 <- c(
  "Strongly Disagree" = "#B2182B",
  "Disagree"          = "#E67E22",
  "Neutral"           = "#F1C40F",
  "Agree"             = "#48C9B0",
  "Strongly Agree"    = "#117A65"
)

# Reshape and clean
df_long_q21 <- student_data_filtered %>%
  pivot_longer(
    cols = starts_with("Q21_"),
    names_to = "Item",
    values_to = "Response"
  ) %>%
  mutate(
    Aspect = recode(Item, !!!aspect_map_q21, .default = NA_character_),
    Response = factor(Response, levels = levels_q21)
  ) %>%
  filter(!is.na(Response), !is.na(Aspect))

#  Calculate percentages
plot_df_q21 <- df_long_q21 %>%
  count(Aspect, Response, name = "n") %>%
  group_by(Aspect) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

# Compute favorable % for sorting
plot_df_q21 <- plot_df_q21 %>%
  group_by(Aspect) %>%
  mutate(pct_fav = sum(pct[Response %in% c("Agree", "Strongly Agree")], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Aspect = fct_reorder(Aspect, pct_fav, .desc = TRUE))

#  Horizontal stacked bar chart
p_q21 <- ggplot(plot_df_q21, aes(y = Aspect, x = pct, fill = Response)) +
  geom_bar(stat = "identity", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = ifelse(pct > 0.04, paste0(round(pct * 100), "%"), "")),
    position = position_stack(vjust = 0.5),
    color = "black", size = 3, fontface = "bold"
  ) +

  # Format axes
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Apply consistent color palette
  scale_fill_manual(values = pal_q21, guide = guide_legend(reverse = TRUE)) +

  # Titles and labels
  labs(
    title = "Q21: Perceived Enhancement of Data Science Workflow with AI",
    subtitle = "Responses grouped by agreement level",
    x = "Percentage of Responses",
    y = "AI-Augmented Workflow Aspect",
    fill = "Response",
    caption = "Note: Percentages represent the proportion of responses within each aspect."
  ) +

  # Apply publication theme
  theme_publication()

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q21_workflow_enhancement_horizontal_pub.png",
  plot = p_q21,
  dpi = 600,
  width = 9,
  height = 6,
  units = "in"
)

print(p_q21)


```

## Q22: What concerns or challenges do you face when integrating AI tools into your data science workflow?

```{r}
# ---- 1) Create binary indicators for each challenge ----
student_data_filtered <- student_data_filtered %>%
  mutate(
    q22_complexity    = str_detect(Q22, regex("complexity", ignore_case = TRUE)),
    q22_accuracy      = str_detect(Q22, regex("accuracy", ignore_case = TRUE)),
    q22_understanding = str_detect(Q22, regex("understanding", ignore_case = TRUE)),
    q22_privacy       = str_detect(Q22, regex("privacy", ignore_case = TRUE)),
    q22_jobimpact     = str_detect(Q22, regex("job", ignore_case = TRUE))
  )

# ---- 2) Summarize counts ----
q22_summary <- tibble(
  Concern = c(
    "Complexity in \nUsing AI Tools",
    "Accuracy & Reliability \nof Predictions",
    "Understanding AI Outputs",
    "Data Security & \nPrivacy Risks",
    "Impact on \nDS Job Roles"
  ),
  Count = c(
    sum(student_data_filtered$q22_complexity, na.rm = TRUE),
    sum(student_data_filtered$q22_accuracy, na.rm = TRUE),
    sum(student_data_filtered$q22_understanding, na.rm = TRUE),
    sum(student_data_filtered$q22_privacy, na.rm = TRUE),
    sum(student_data_filtered$q22_jobimpact, na.rm = TRUE)
  )
)

# ---- 3) Add percentages ----
n22 = sum(!is.na(student_data_filtered$Q22))  # update this with your actual N
q22_summary <- q22_summary %>%
  #mutate(Percent = Count / total_n)
mutate(Percent = Count / n22)

# ---- 4) Horizontal bar chart with refined theme ----
p_q22 <- ggplot(q22_summary, aes(y = fct_reorder(Concern, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percent labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # X-axis formatting
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(q22_summary$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles & caption
  labs(
    title = "Q22: Challenges in Integrating AI Tools into Data Science Workflows",
    subtitle = "Key challenges identified from open-ended responses",
    x = "Percentage of Students",
    y = "Challenge",
    caption = "Note: Multiple responses were allowed."
  ) +

  # Apply your publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# ---- 5) Save as high-resolution PNG ----
ggsave(
  filename = "Student_figures/Q22_challenges_workflow_pub.png",
  plot = p_q22,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q22)

```

## Q23: What concerns do you have about integrating AI tools into education in general?

```{r}
# binary indicators
student_data_filtered <- student_data_filtered %>%
  mutate(
    q23_undermines_value = str_detect(Q23, regex("undermine|value", ignore_case = TRUE)),
    q23_limits_social    = str_detect(Q23, regex("interact|social", ignore_case = TRUE)),
    q23_hinders_skills   = str_detect(Q23, regex("skills|teamwork|problem", ignore_case = TRUE)),
    q23_overreliance     = str_detect(Q23, regex("over[- ]?reliant", ignore_case = TRUE))
  )

# Summarize counts
q23_summary <- tibble(
  Concern = c(
    "Undermines Value of \nUniversity Education",
    "Limits Social \nInteraction",
    "Hinders Development \nof Transferable Skills",
    "Over-Reliance on \nGenerative AI"
  ),
  Count = c(
    sum(student_data_filtered$q23_undermines_value, na.rm = TRUE),
    sum(student_data_filtered$q23_limits_social, na.rm = TRUE),
    sum(student_data_filtered$q23_hinders_skills, na.rm = TRUE),
    sum(student_data_filtered$q23_overreliance, na.rm = TRUE)
  )
)

#  percentages
total_n <- 120
q23_summary <- q23_summary %>%
  mutate(Percent = Count / total_n)

# ---- 4) Horizontal bar chart with refined styling ----
p_q23 <- ggplot(q23_summary, aes(y = fct_reorder(Concern, Percent), x = Percent)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +

  # Inline percentage labels
  geom_text(
    aes(label = paste0(round(Percent * 100, 1), "%")),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +

  # Format axes
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(q23_summary$Percent) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +

  # Titles, labels, caption
  labs(
    title = "Q23: Concerns About Integrating AI into Education",
    subtitle = "Top concerns identified from open-ended responses",
    x = "Percentage of Students",
    y = "Concern",
    caption = "Note: Multiple responses were allowed."

  ) +

  # Apply publication theme
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

#  high-resolution PNG
ggsave(
  filename = "Student_figures/Q23_general_concerns_horizontal_pub.png",
  plot = p_q23,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q23)


```

## Q24: Do you believe that AI tools should complement traditional methods in data science, or replace them?
```{r}
#  Wrap labels
student_data_filtered$Q24 <- str_wrap(student_data_filtered$Q24, width = 35)

# Tabulate and compute percentages
plot_df_q24 <- student_data_filtered %>%
  filter(!is.na(Q24)) %>%
  count(Q24, name = "n") %>%
  mutate(pct = n / total_n)

# Horizontal bar chart
p_q24 <- ggplot(plot_df_q24, aes(y = fct_reorder(Q24, pct), x = pct)) +
  geom_bar(stat = "identity", fill = "#48C9B0", color = "gray30", width = 0.7) +
  geom_text(
    aes(label = percent(pct, accuracy = 0.1)),
    hjust = -0.15, fontface = "bold", size = 3.6, color = "black"
  ) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, max(plot_df_q24$pct) + 0.08),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title = "Q24: Should AI Tools Complement or Replace Traditional Methods?",
    subtitle = "Distribution of open-ended responses",
    x = "Percentage of Valid Responses",
    y = "Response",
    caption = "Note: Percentages represent non-missing responses."
  ) +
  theme_publication() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10.5, lineheight = 1.1),
    plot.margin = margin(10, 20, 25, 15)
  )

# high-resolution PNG
ggsave(
  filename = "Student_figures/Q24_complement_or_replace_horizontal_pub.png",
  plot = p_q24,
  dpi = 600,
  width = 8.5,
  height = 5.5,
  units = "in"
)

print(p_q24)

```

#cfa

```{r cfa_data_refresh, message=FALSE, warning=FALSE}
# Refresh the analysis objects using the same repo-relative loader
student_data_path <- find_student_file()
student_import <- load_student_data(student_data_path)
student_header_preview <- student_import$student_header_preview
student_data_raw <- student_import$student_data_raw
student_data_filtered <- student_import$student_data_filtered
student_header_labels <- student_import$student_header_labels

cat("Remaining sample size after quality filters:", nrow(student_data_filtered), "students\n")
dim(student_data_filtered)
```

```{r }


questions <- student_data_filtered |>
  select(
    starts_with("Q10_"),
    starts_with("Q12_"),
    starts_with("Q18_"),
    starts_with("Q15_"),
    starts_with("Q21_")
  ) |>
  select(
    Q10_1, Q10_2, Q10_3, Q10_4, Q10_5,
    Q12_1, Q12_2, Q12_3, Q12_4,
    Q18_1, Q18_2, Q18_3, Q18_4, Q18_5,
    Q21_1, Q21_2, Q21_3, Q21_4, Q21_5,
    Q15_4
  )


missing_summary <- questions |>
  pivot_longer(
    cols = everything(),
    names_to = "Question",
    values_to = "Response"
  ) |>
  group_by(Question) |>
  summarise(
  Valid_Responses = sum(!is.na(Response)),
  Missing_Count = sum(is.na(Response)),
  Percent_Missing = round(100 * Missing_Count / n(), 2),
  Percent_Not_Missing = round(100 * Valid_Responses / n(), 2),
  .groups = "drop"
)







  missing_summary

```


## Familiarity Score Q10

# How familiar are you with the following AI technologies in data science?


**CLASS GROUPS AND RECODING **


```{r}

# Class Groups

student_data_filtered <- student_data_filtered %>%
  mutate(
    Q5 = replace_na(Q5, "Graduate student"),
    Q5 = factor(Q5, levels = c("Freshman", "Sophomore", "Junior", "Senior", "Graduate student")))

student_data_filtered <- student_data_filtered |>
  mutate(
    class_group = case_when(
      Q5 %in% c("Freshman", "Sophomore") ~ "Lower Division",
      Q5 %in% c("Junior", "Senior")      ~ "Upper Division",
      Q5 == "Graduate student"           ~ "Graduate",
      TRUE                               ~ NA_character_
    )
  )


student_data_filtered <- student_data_filtered |>
  mutate(  class_group = factor(class_group, levels = c("Lower Division","Upper Division", "Graduate")))

# Recoding to numeric


student_data_filtered <- student_data_filtered |>
  mutate(Q10_1 =
           recode(Q10_1,
                  "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5),
        Q10_2 =
              recode(Q10_2,
                "Not at All Familiar" = 1,
                "Not Very Familiar" = 2,
                "Neutral" = 3,
                "Somewhat Familiar" = 4,
                "Very Familiar" = 5),
        Q10_3 =
         recode(Q10_3,
                  "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5),
         Q10_4 =
         recode(Q10_4,
                "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5),
        Q10_5 =
         recode(Q10_5,
                "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5))


head(student_data_filtered)


```


```{r}


student_data_filtered <- student_data_filtered |>
  rowwise() |>
  mutate(
    Q10_comp = ifelse(
      all(is.na(c_across(Q10_1:Q10_5))),
      NA_real_,
      mean(c_across(Q10_1:Q10_5), na.rm = TRUE)
    )
  ) |> ungroup()

head(student_data_filtered)


```


# SUMMARY


```{r}

Q10_Sum <- student_data_filtered |>
  summarise(n = sum(!is.na(Q10_comp)),
            Min = min(Q10_comp, na.rm = TRUE),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE),
            Max = max(Q10_comp, na.rm = TRUE))

print(Q10_Sum)


```


# SUMMARY BY GENDER AND CLASS


```{r}
# by gender

student_data_filtered <- student_data_filtered %>%
  filter(Q3 %in% c("Female", "Male"))

Q10_bygender <- student_data_filtered |>
  group_by(Q3) |>
   filter(!is.na(class_group)) |>
  summarise(n= sum(!is.na(Q10_comp)),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE))

print(Q10_bygender)


# by class

Q10_class <- student_data_filtered |>
  group_by(class_group) |>
  filter(!is.na(class_group)) |>
  summarise(n= sum(!is.na(Q10_comp)),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE))

print(Q10_class)


```

# BOXPLOT  FOR CLASS AND GENDER


```{r, warning=FALSE, message=FALSE}

student_data_filtered |>
  ggplot(aes(x = Q3, y = Q10_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = FALSE) +
  labs(
    title = "Q10 Composite Score by Gender",
    x = "Gender",
    y = "Q10 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )


student_data_filtered |>
  ggplot(aes(x = class_group, y = Q10_comp )) +
  geom_boxplot(width = 0.6, alpha = 0.8) +
  labs(
    title = "Q10 Composite Score by Class Group",
    x = "Class Group",
    y = "Q10 Composite Score"
  )

student_data_filtered |> filter(!is.na(class_group), !is.na(Q10_comp))|>
 ggplot(aes(x = class_group, y = Q10_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = TRUE) +
  labs(
    title = "Q10 Composite Score by Class Group and Gender",
    x = "Class Group",
    y = "Q10 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )


```


## Awareness of limitations Score Q18


# Q18 - To what extent are you aware of the following limitations of using AI for data science tasks?


```{r}

# Recoding to numeric


student_data_filtered <- student_data_filtered |>
  mutate(Q18_1 =
           recode(Q18_1,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
        Q18_2 =
              recode(Q18_2,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
        Q18_3 =
         recode(Q18_3,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
         Q18_4 =
         recode(Q18_4,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
        Q18_5 =
         recode(Q18_5,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5))


student_data_filtered <- student_data_filtered |>
  mutate(Q15_1 =
           recode(Q15_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q15_2 =
          recode(Q15_2,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),

        Q15_3 =
         recode(Q15_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
         Q15_4 =    # reverse coding
         recode(Q15_4,
                  "Strongly Disagree" = 5,
                  "Disagree" = 4,
                  "Neutral" = 3,
                  "Agree" = 2,
                  "Strongly Agree" = 1))


head(student_data_filtered)


```


```{r}


student_data_filtered <- student_data_filtered |>
  rowwise() |>
  mutate(
    Q18_comp = ifelse(
      all(is.na(c_across(c(Q18_1:Q18_5)))),
      NA_real_,
      mean(c_across(c(Q18_1:Q18_5)), na.rm = TRUE)
    )
  ) |>
  ungroup()

head(student_data_filtered)


```


# summary

```{r}

Q18_Sum <- student_data_filtered |>
  summarise(n = sum(!is.na(Q18_comp)),
            Min = min(Q18_comp, na.rm = TRUE),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE),
            Max = max(Q18_comp, na.rm = TRUE))

print(Q18_Sum)

```


# SUMMARY BY GENDER AND CLASS


```{r}

# by gender

Q18_bygender <- student_data_filtered |>
  group_by(Q3) |>
   filter(!is.na(class_group), !is.na(Q3)) |>
  summarise(n = sum(!is.na(Q18_comp)),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE))

print(Q18_bygender)


# by class

Q18_class <- student_data_filtered |>
  group_by(class_group) |>
  filter(!is.na(class_group)) |>
  summarise(n = sum(!is.na(Q18_comp)),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE))

print(Q18_class)


```


# BOXPLOT  FOR CLASS AND GENDER


```{r, warning=FALSE,message=FALSE}

student_data_filtered |>
  ggplot(aes(x = Q3, y = Q18_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = FALSE) +
  labs(
    title = "Q18 Composite Score by Gender",
    x = "Gender",
    y = "Q18 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )


student_data_filtered |>
  ggplot(aes(x = class_group, y = Q18_comp )) +
  geom_boxplot(width = 0.6, alpha = 0.8) +
  labs(
    title = "Q18 Composite Score by Class Group",
    x = "Class Group",
    y = "Q18 Composite Score"
  )

student_data_filtered |> filter(!is.na(class_group), !is.na(Q18_comp))|>
 ggplot(aes(x = class_group, y = Q18_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = TRUE) +
  labs(
    title = "Q18 Composite Score by Class Group and Gender",
    x = "Class Group",
    y = "Q18 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )


```


## Perceptions Q12 AND Q21

# Q12 - To what extent do you agree with the following statements regarding AI tools in data science?

# Q21 - Do you think AI tools can significantly enhance the following aspects of your data science workflow?


# RECODING


```{r, warning=FALSE, message= FALSE}

student_data_filtered <- student_data_filtered |>
  mutate(Q12_1 =
           recode(Q12_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q12_2 =
          recode(Q12_2,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),

        Q12_3 =
         recode(Q12_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
         Q12_4 =  # reverse coding
         recode(Q12_4,
                  "Strongly Disagree" = 5,
                  "Disagree" = 4,
                  "Neutral" = 3,
                  "Agree" = 2,
                  "Strongly Agree" = 1))


student_data_filtered <- student_data_filtered |>
  mutate(Q21_1 =
           recode(Q21_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q21_2 =
              recode(Q21_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q21_3 =
         recode(Q21_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
         Q21_4 =
         recode(Q21_4,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q21_5 =
         recode(Q21_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5))


head(student_data_filtered)

```


```{r}


student_data_filtered <- student_data_filtered |>
  rowwise() |>
  mutate(
    Qper_comp = ifelse(
      all(is.na(c_across(c(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4)))),
      NA_real_,
      mean(c_across(c(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4)), na.rm = TRUE)
    )
  ) |>
  ungroup()

head(student_data_filtered)


```


# Summary


```{r}

Qper_Sum <- student_data_filtered |>
  summarise(n = sum(!is.na(Qper_comp)),
            Min = min(Qper_comp, na.rm = TRUE),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE),
            Max = max(Qper_comp, na.rm = TRUE))

print(Qper_Sum)

```


# SUMMARY BY GENDER AND CLASS

```{r}

# by gender

Qper_bygender <- student_data_filtered |>
  group_by(Q3) |>
   filter(!is.na(class_group), !is.na(Q3)) |>
  summarise(n = sum(!is.na(Qper_comp)),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE))

print(Qper_bygender)


# by class

Qper_class <- student_data_filtered |>
  group_by(class_group) |>
  filter(!is.na(class_group)) |>
  summarise(n = sum(!is.na(Qper_comp)),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE))

print(Qper_class)


```


# BOXPLOT  FOR CLASS AND GENDER


```{r, warning=FALSE, message=FALSE}

student_data_filtered |>
  ggplot(aes(x = Q3, y = Qper_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = FALSE) +
  labs(
    title = "Q12 and Q21 Composite Score by Gender",
    x = "Gender",
    y = "Q12 and Q21 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )


student_data_filtered |>
  ggplot(aes(x = class_group, y = Qper_comp )) +
  geom_boxplot(width = 0.6, alpha = 0.8) +
  labs(
    title = "Q12 and Q21 Composite Score by Class Group",
    x = "Class Group",
    y = "Q12 and Q21 Composite Score"
  )

student_data_filtered |> filter(!is.na(class_group), !is.na(Qper_comp))|>
 ggplot(aes(x = class_group, y = Qper_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = TRUE) +
  labs(
    title = "Q12 and Q21 Composite Score by Class Group and Gender",
    x = "Class Group",
    y = "Q12 and Q21 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )


```


### Factor Analysis


##  EFA AND CFA Familiarity Score Q10


# Exploratory Factor Analysis


```{r}

Q10_fam <- student_data_filtered |> select(Q10_1:Q10_5) |>
  na.omit()

factanal(Q10_fam, factors = 1)


```


## Confirmatory Factor Analysis (CFA)


```{r}

Q10_model <- '
  Q10_factor =~ Q10_1 + Q10_2 + Q10_3 + Q10_4 + Q10_5
'


Q10_CFA <- cfa(model = Q10_model, data = student_data_filtered)


summary(Q10_CFA, fit.measures = TRUE, standardized = TRUE)


```

```{r}

Alpha <- psych::alpha(student_data_filtered |> dplyr::select(Q10_1:Q10_5))


Alpha$total


```

```{r}

fam_table  <- data.frame(
  Ques_Type = ("Familiarity (Q10)"),
  Chi_Square = fitMeasures(Q10_CFA, "chisq"),
  df = fitMeasures(Q10_CFA,"df"),
  P_value = fitMeasures(Q10_CFA, "pvalue"),
  CFI = fitMeasures(Q10_CFA, "cfi"),
  TLI = fitMeasures(Q10_CFA, "tli"),
  RMSEA = fitMeasures(Q10_CFA, "rmsea"),
  Raw_Alpha = Alpha$total$raw_alpha,
  Std_Alpha = Alpha$total$std.alpha,
  row.names = NULL
)


fam_table

```


## EFA AND CFA for Awareness of limitations Score Q18 and  Q15_4


# Exploratory Factor Analysis


```{r}

Q_awa <- student_data_filtered |> select(Q18_1:Q18_5) |>
  na.omit()


factanal(Q_awa, factors = 1)


```


# Confirmatory Factor Analysis (CFA)


```{r}

Q_awa_model <- '
  Qawa_factor =~ Q18_1 + Q18_2 + Q18_3 + Q18_4 + Q18_5
'

Q_awa_cfa <- cfa(model = Q_awa_model, data = student_data_filtered)

summary(Q_awa_cfa, fit.measures = TRUE, standardized = TRUE)


```


```{r}
Alpha1  <- psych::alpha(student_data_filtered |> dplyr::select(Q18_1:Q18_5))


Alpha1$total
```


```{r}
awa_table  <- data.frame(
  Ques_Type = ("Awar of lim (Q18)"),
  Chi_Square = fitMeasures(Q_awa_cfa, "chisq"),
  df = fitMeasures(Q_awa_cfa,"df"),
  P_value = fitMeasures(Q_awa_cfa, "pvalue"),
  CFI = fitMeasures(Q_awa_cfa, "cfi"),
  TLI = fitMeasures(Q_awa_cfa, "tli"),
  RMSEA = fitMeasures(Q_awa_cfa, "rmsea"),
  Raw_Alpha = Alpha1$total$raw_alpha,
  Std_Alpha = Alpha1$total$std.alpha,
  row.names = NULL
)


awa_table
```


## EFA AND CFA for Perceptions Q12 AND Q21

# Exploratory Factor Analysis


```{r}

Q_per <- student_data_filtered |> select(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4) |>
  na.omit()


factanal(Q_per,factors = 1)


```


# Confirmatory Factor Analysis (CFA)


```{r}

Q_per_model <- '
  Q12_factor =~ Q12_1 + Q12_2 + Q21_1 + Q21_2 + Q21_4
'

Q_per_cfa <- cfa(model = Q_per_model, data = student_data_filtered)

summary(Q_per_cfa, fit.measures = TRUE, standardized = TRUE)


```


```{r}

Alpha2 <- psych::alpha(
  student_data_filtered |> dplyr::select(Q12_1:Q12_2, Q21_1:Q21_2, Q21_4)
)

Alpha2$total
```


# Table

```{r}

per_table  <- data.frame(
  Ques_Type = ("Perception (Q12,Q21)"),
  Chi_Square = fitMeasures(Q_per_cfa, "chisq"),
  df = fitMeasures(Q_per_cfa,"df"),
  P_value = fitMeasures(Q_per_cfa, "pvalue"),
  CFI = fitMeasures(Q_per_cfa, "cfi"),
  TLI = fitMeasures(Q_per_cfa, "tli"),
  RMSEA = fitMeasures(Q_per_cfa, "rmsea"),
  Raw_Alpha = Alpha2$total$raw_alpha,
  Std_Alpha = Alpha2$total$std.alpha,
  row.names = NULL
)


per_table


```


# Confirmatory Factor Analysis Results

```{r}

Summary_Table <- bind_rows(fam_table,awa_table,per_table)

Summary_Table

```


