---
output:
  html_document: default
  pdf_document: default
---

# Reproducible analysis of the student composite score analysis


```{r setup-student-analysis, warning=FALSE, message=FALSE}
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
  dplyr, janitor, ggplot2, rlang, ggmosaic, knitr, forcats, tidyr,
  stringr, psych, readxl, scales, kableExtra, grid, tibble, lavaan, infer
)

# Create output directory for saved figures
if (!dir.exists("Student_figures")) dir.create("Student_figures", recursive = TRUE)

find_student_file <- function() {
  candidates <- c(
    "AI4DS_Student_Survey_October_5_ 2025.xlsx",
    file.path("data", "AI4DS_Student_Survey_October_5_ 2025.xlsx"),
    # Allows knitting from the RCode/ directory in the repository.
    file.path("..", "data", "AI4DS_Student_Survey_October_5_ 2025.xlsx")
  )
  found <- candidates[file.exists(candidates)]
  if (length(found) == 0) {
    stop(
      "Student survey Excel file not found. ",
      "Expected it in the repo root, data/, or ../data/ relative to the working directory. ",
      "Update `find_student_file()` only if the repository structure has changed."
    )
  }
  found[[1]]
}

load_student_data <- function(path = find_student_file(), min_duration = 120, min_progress = 25) {
  raw_preview <- readxl::read_excel(path)
  short <- readxl::read_excel(path, range = "A1:BN1", col_names = FALSE) |> unlist() |> as.character()
  long  <- readxl::read_excel(path, range = "A2:BN2", col_names = FALSE) |> unlist() |> as.character()
  n_cols <- ncol(raw_preview)
  short <- short[seq_len(n_cols)]
  long  <- long[seq_len(n_cols)]
  student_df <- readxl::read_excel(path, skip = 2, col_names = short)
  student_header_labels <- setNames(long, short)
  filtered_df <- student_df |>
    dplyr::filter(`Duration (in seconds)` >= min_duration, Progress >= min_progress)
  list(student_header_preview = raw_preview, student_data_raw = student_df, student_data_filtered = filtered_df, student_header_labels = student_header_labels)
}
```

```{r import-student-data, warning=FALSE, message=FALSE}
student_data_path <- find_student_file()
student_import <- load_student_data(student_data_path)
student_header_preview <- student_import$student_header_preview
student_data_raw <- student_import$student_data_raw
student_data_filtered <- student_import$student_data_filtered
student_header_labels <- student_import$student_header_labels

cat("Remaining sample size after quality filters:", nrow(student_data_filtered), "students\n")

dim(student_data_filtered)
```



```{r}

library(dplyr)
library(tidyr)


questions <- student_data_filtered |>
  select(
    starts_with("Q10_"),
    starts_with("Q12_"),
    starts_with("Q18_"),
    starts_with("Q15_"),
    starts_with("Q21_")
  ) |>
  select(
    Q10_1, Q10_2, Q10_3, Q10_4, Q10_5,
    Q12_1, Q12_2, Q12_3, Q12_4,
    Q18_1, Q18_2, Q18_3, Q18_4, Q18_5,
    Q21_1, Q21_2, Q21_3, Q21_4, Q21_5,
    Q15_4
  )


missing_summary <- questions |>
  pivot_longer(
    cols = everything(),
    names_to = "Question",
    values_to = "Response"
  ) |>
  group_by(Question) |>
  summarise(
  Valid_Responses = sum(!is.na(Response)),
  Missing_Count = sum(is.na(Response)),
  Percent_Missing = round(100 * Missing_Count / n(), 2),
  Percent_Not_Missing = round(100 * Valid_Responses / n(), 2),
  .groups = "drop"
)


  missing_summary


```


# Familiarity Score Q10

## How familiar are you with the following AI technologies in data science?


**CLASS GROUPS AND RECODING**


```{r CLASS GROUPS AND RECODING}

# Class Groups
# In the initial survey version, "Graduate Student" was not available  as a Q5 response option. Review of the reported-major field established that respondents with missing Q5 values in this portion  of the data were graduate students. The analysis therefore recoded #these Q5 missing values as Graduate Student;


student_data_filtered <- student_data_filtered %>%
  mutate(
    Q5 = replace_na(Q5, "Graduate Student"),
    Q5 = factor(Q5, levels = c("Freshman", "Sophomore", "Junior", "Senior", "Graduate Student")))

student_data_filtered <- student_data_filtered |>
  mutate(
    class_group = case_when(
      Q5 %in% c("Freshman", "Sophomore") ~ "Lower Division",
      Q5 %in% c("Junior", "Senior")      ~ "Upper Division",
      Q5 == "Graduate Student"           ~ "Graduate",
      TRUE                               ~ NA_character_
    )
  )


student_data_filtered <- student_data_filtered |>
  mutate(  class_group = factor(class_group, levels = c("Lower Division","Upper Division", "Graduate")))

# Recoding to numeric


student_data_filtered <- student_data_filtered |>
  mutate(Q10_1 =
           recode(Q10_1,
                  "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5),
        Q10_2 = 
              recode(Q10_2,
                "Not at All Familiar" = 1,
                "Not Very Familiar" = 2,
                "Neutral" = 3,
                "Somewhat Familiar" = 4,
                "Very Familiar" = 5),
        Q10_3 =
         recode(Q10_3,
                  "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5),
         Q10_4 = 
         recode(Q10_4,
                  "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5),
        Q10_5 =
         recode(Q10_5,
                "Not at All Familiar" = 1,
                  "Not Very Familiar" = 2,
                  "Neutral" = 3,
                  "Somewhat Familiar" = 4,
                  "Very Familiar" = 5))



head(student_data_filtered)


```






```{r}

student_data_filtered <- student_data_filtered |>
  rowwise() |>
  mutate(
    Q10_comp = ifelse(
      all(is.na(c_across(Q10_1:Q10_5))),
      NA_real_,
      mean(c_across(Q10_1:Q10_5), na.rm = TRUE)
    )
  ) |> ungroup()

head(student_data_filtered)


```







# Awareness of limitations Score Q18


## Q18 - To what extent are you aware of the following limitations of using AI for data science tasks?


## RECODING


```{r Awareness of limitations Score Q18}

# Recoding to numeric

student_data_filtered <- student_data_filtered |>
  mutate(Q18_1 =
           recode(Q18_1,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
        Q18_2 = 
              recode(Q18_2,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
        Q18_3 =
         recode(Q18_3,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
         Q18_4 = 
         recode(Q18_4,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5),
        Q18_5 =
         recode(Q18_5,
                  "Not at All Aware" = 1,
                  "Not Very Aware" = 2,
                  "Neutral" = 3,
                  "Somewhat Aware" = 4,
                  "Very Aware" = 5))



student_data_filtered <- student_data_filtered |>
  mutate(Q15_1 =
           recode(Q15_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q15_2 = 
          recode(Q15_2,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
              
        Q15_3 =
         recode(Q15_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
         Q15_4 =    # reverse coding
         recode(Q15_4,
                  "Strongly Disagree" = 5,
                  "Disagree" = 4,
                  "Neutral" = 3,
                  "Agree" = 2,
                  "Strongly Agree" = 1))



head(student_data_filtered)



```



```{r}

student_data_filtered <- student_data_filtered |>
  rowwise() |>
  mutate(
    Q18_comp = ifelse(
      all(is.na(c_across(c(Q18_1:Q18_5)))),
      NA_real_,
      mean(c_across(c(Q18_1:Q18_5)), na.rm = TRUE)
    )
  ) |>
  ungroup()

head(student_data_filtered)


```



## summary 

```{r}

Q18_Sum <- student_data_filtered |>
  summarise(n = sum(!is.na(Q18_comp)),
            Min = min(Q18_comp, na.rm = TRUE),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE),
            Max = max(Q18_comp, na.rm = TRUE))

print(Q18_Sum)

```




## SUMMARY BY GENDER AND CLASS


```{r Awareness of Limitations by Gender & Class level}

# by gender

Q18_bygender <- student_data_filtered |> 
  group_by(Q3) |>
   filter(!is.na(class_group), !is.na(Q3)) |>
  summarise(n = sum(!is.na(Q18_comp)),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE)) 

print(Q18_bygender)




# by class

Q18_class <- student_data_filtered |> 
  group_by(class_group) |>
  filter(!is.na(class_group)) |>
  summarise(n = sum(!is.na(Q18_comp)),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE))

print(Q18_class)


```




## BOXPLOT  FOR CLASS AND GENDER


```{r, warning=FALSE,message=FALSE}

student_data_filtered |>
  ggplot(aes(x = Q3, y = Q18_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = FALSE) +
  labs(
    title = "Q18 Composite Score by Gender",
    x = "Gender",
    y = "Q18 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )



student_data_filtered |>
  ggplot(aes(x = class_group, y = Q18_comp )) +
  geom_boxplot(width = 0.6, alpha = 0.8) +
  labs(
    title = "Q18 Composite Score by Class Group",
    x = "Class Group",
    y = "Q18 Composite Score"
  )

student_data_filtered |> filter(!is.na(class_group), !is.na(Q18_comp))|>
 ggplot(aes(x = class_group, y = Q18_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = TRUE) +
  labs(
    title = "Awareness of Limitations Scores by Class & Gender",
    x = "Class Group",
    y = "Limitations Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )


```

# Perceptions Q12 AND Q21

## Q12 - To what extent do you agree with the following statements regarding AI tools in data science? 

## Q21 - Do you think AI tools can significantly enhance the following aspects of your data science workflow?  


## RECODING



```{r Perceptions Score}

student_data_filtered <- student_data_filtered |>
  mutate(Q12_1 =
           recode(Q12_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q12_2 = 
          recode(Q12_2,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
              
        Q12_3 =
         recode(Q12_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
         Q12_4 =  # reverse coding
         recode(Q12_4,
                  "Strongly Disagree" = 5,
                  "Disagree" = 4,
                  "Neutral" = 3,
                  "Agree" = 2,
                  "Strongly Agree" = 1))




student_data_filtered <- student_data_filtered |>
  mutate(Q21_1 =
           recode(Q21_1,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q21_2 = 
              recode(Q21_2,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
        Q21_3 =
         recode(Q21_3,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),
         Q21_4 = 
         recode(Q21_4,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5),

        
        Q21_5 =
         recode(Q21_5,
                  "Strongly Disagree" = 1,
                  "Disagree" = 2,
                  "Neutral" = 3,
                  "Agree" = 4,
                  "Strongly Agree" = 5))

head(student_data_filtered)

```






```{r}


# compute the arithmetic mean across available selected items for each respondent; return NA only when all five
# selected items are missing.
student_data_filtered <- student_data_filtered |>
  rowwise() |>
  mutate(
    Qper_comp = ifelse(
      all(is.na(c_across(c(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4)))),
      NA_real_,
      mean(c_across(c(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4)), na.rm = TRUE)
    )
  ) |>
  ungroup()

head(student_data_filtered)




```



## Summary


```{r}

Qper_Sum <- student_data_filtered |>
  summarise(n = sum(!is.na(Qper_comp)),
            Min = min(Qper_comp, na.rm = TRUE),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE),
            Max = max(Qper_comp, na.rm = TRUE))

print(Qper_Sum)

```



## SUMMARY BY GENDER AND CLASS

```{r Preception Scores by Gender & Class}

# by gender

Qper_bygender <- student_data_filtered |> 
  group_by(Q3) |>
   filter(!is.na(class_group), !is.na(Q3)) |>
  summarise(n = sum(!is.na(Qper_comp)),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE)) 

print(Qper_bygender)




# by class

Qper_class <- student_data_filtered |> 
  group_by(class_group) |>
  filter(!is.na(class_group)) |>
  summarise(n = sum(!is.na(Qper_comp)),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE))

print(Qper_class)


```






## BOXPLOT  FOR CLASS AND GENDER


```{r, warning=FALSE, message=FALSE}

student_data_filtered |>
  ggplot(aes(x = Q3, y = Qper_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = FALSE) +
  labs(
    title = "Perception Composite Score by Gender",
    x = "Gender",
    y = "Perception Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )



student_data_filtered |>
  ggplot(aes(x = class_group, y = Qper_comp )) +
  geom_boxplot(width = 0.6, alpha = 0.8) +
  labs(
    title = "Perception Composite Score by Class Group",
    x = "Class Group",
    y = "Perception Score"
  )

student_data_filtered |> filter(!is.na(class_group), !is.na(Qper_comp))|>
 ggplot(aes(x = class_group, y = Qper_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = TRUE) +
  labs(
    title = "Perception Composite Score by Class Group and Gender",
    x = "Class Group",
    y = "Perception Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )


```



# Overall Group Comparison Analyses

## Independent Samples T-Tests for Gender Differences in Composite Scores


```{r}
library(infer)

   #  Familiarity
student_data_filtered |>
  t_test(Q10_comp ~ Q3, order = c("Male", "Female") )

  #   Awareness of limitations
student_data_filtered |>
  t_test(Q18_comp ~ Q3, order = c("Male", "Female") )


  #   Perceptions
student_data_filtered |>
  t_test(Qper_comp ~ Q3, order = c("Male", "Female") )
```

## ANOVA and Tukey Multiple Comparisons for Class Group Differences in All Composite Scores


```{r}
   #  Familiarity
 aov_10 <-  aov(Q10_comp ~ class_group, data = student_data_filtered)

summary(aov_10)
TukeyHSD(aov_10)

  #   Awareness of limitations
aov_lim <-   aov(Q18_comp ~ class_group, data = student_data_filtered  )
summary(aov_lim)
TukeyHSD(aov_lim)

  #   Perceptions
aov_per <-   aov(Qper_comp ~ class_group, data = student_data_filtered  )
summary(aov_per)
TukeyHSD(aov_per)

```

## T-Tests Comparing Graduate and Undergraduate Students on All Composite Scores


```{r}
student_data_filtered <- student_data_filtered |> mutate(class_group2 = ifelse(class_group == "Graduate" ,
                                           "Graduate", "Undergraduate"))

# familiarity

student_data_filtered |>
t_test(Q10_comp ~ class_group2, order = c("Graduate", "Undergraduate") )


 # Awareness of Limitations

student_data_filtered |>
  t_test(Q18_comp ~ class_group2, order = c("Graduate", "Undergraduate") )

# Perceptions

student_data_filtered |>
  t_test(Qper_comp ~ class_group2, order = c("Graduate", "Undergraduate") )
```


## Summary Statistics (n, Mean, SD) by Class Group2 (Graduate vs. Undergraduate)
## Not included in the manuscript.

```{r}
  # familiarity

Q10_byclass2 <- student_data_filtered |> 
  group_by(class_group2) |>
   filter(!is.na(class_group2)) |>
  summarise(n= sum(!is.na(Q10_comp)),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE)) 

print(Q10_byclass2)

  # Awareness of Limitations

Q18_byclass2 <- student_data_filtered |> 
  group_by(class_group2) |>
   filter(!is.na(class_group2)) |>
  summarise(n = sum(!is.na(Q18_comp)),
            Mean = mean(Q18_comp, na.rm = TRUE),
            SD = sd(Q18_comp, na.rm =TRUE)) 

print(Q18_byclass2)



  # Perceptions

Qper_byclass2 <- student_data_filtered |> 
  group_by(class_group2) |>
   filter(!is.na(class_group2)) |>
  summarise(n = sum(!is.na(Qper_comp)),
            Mean = mean(Qper_comp, na.rm = TRUE),
            SD = sd(Qper_comp, na.rm =TRUE)) 

print(Qper_byclass2)


```


## Confirmatory Factor Analysis (CFA)


```{r}

library(lavaan)

Q10_model <- 'Q10_factor =~ Q10_1 + Q10_2 + Q10_3 + Q10_4 + Q10_5'

Q10_CFA <- cfa(model = Q10_model, data = student_data_filtered)


summary(Q10_CFA, fit.measures = TRUE, standardized = TRUE)




```

```{r}

Alpha <- psych::alpha(
  student_data_filtered |>
    dplyr::select(Q10_1:Q10_5)
)


Alpha$total


```




```{r}

fam_table  <- data.frame(
  Ques_Type = ("Familiarity (Q10)"),
  Chi_Square = fitMeasures(Q10_CFA, "chisq"),
  df = fitMeasures(Q10_CFA,"df"),
  P_value = fitMeasures(Q10_CFA, "pvalue"),
  CFI = fitMeasures(Q10_CFA, "cfi"),
  TLI = fitMeasures(Q10_CFA, "tli"),
  RMSEA = fitMeasures(Q10_CFA, "rmsea"),
  Raw_Alpha = Alpha$total$raw_alpha,
  Std_Alpha = Alpha$total$std.alpha,
  row.names = NULL
)


fam_table

```


## EFA AND CFA for Awareness of limitations Score Q18 and  Q15_4
### Exploratory Factor Analysis 


```{r}

Q_awa <- student_data_filtered |> select(Q18_1:Q18_5) |>
  na.omit()



factanal(Q_awa, factors = 1)



```


## Confirmatory Factor Analysis (CFA)


```{r}

Q_awa_model <- '
  Qawa_factor =~ Q18_1 + Q18_2 + Q18_3 + Q18_4 + Q18_5
'

Q_awa_cfa <- cfa(model = Q_awa_model, data = student_data_filtered)

summary(Q_awa_cfa, fit.measures = TRUE, standardized = TRUE)


```


```{r}

Alpha1 <- psych::alpha(
  student_data_filtered |>
    dplyr::select(Q18_1:Q18_5)
)


Alpha1$total
```




```{r}
awa_table  <- data.frame(
  Ques_Type = ("Awar of lim (Q18)"),
  Chi_Square = fitMeasures(Q_awa_cfa, "chisq"),
  df = fitMeasures(Q_awa_cfa,"df"),
  P_value = fitMeasures(Q_awa_cfa, "pvalue"),
  CFI = fitMeasures(Q_awa_cfa, "cfi"),
  TLI = fitMeasures(Q_awa_cfa, "tli"),
  RMSEA = fitMeasures(Q_awa_cfa, "rmsea"),
  Raw_Alpha = Alpha1$total$raw_alpha,
  Std_Alpha = Alpha1$total$std.alpha,
  row.names = NULL
)


awa_table
```



## EFA AND CFA for Perceptions Q12 AND Q21
### Exploratory Factor Analysis 



```{r Perceptions}

Q_per <- student_data_filtered |> select(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4) |>
  na.omit()



factanal(Q_per,factors = 1)


```

## Confirmatory Factor Analysis (CFA)


```{r}

Q_per_model <- '
  Q12_factor =~ Q12_1 + Q12_2 + Q21_1 + Q21_2 + Q21_4 
'

Q_per_cfa <- cfa(model = Q_per_model, data = student_data_filtered)

summary(Q_per_cfa, fit.measures = TRUE, standardized = TRUE)



```



```{r}

Alpha2 <- psych::alpha(
  student_data_filtered |>
    dplyr::select(Q12_1:Q12_2, Q21_1:Q21_2,Q21_4)
)

Alpha2$total

```





## Table of 

```{r}

per_table  <- data.frame(
  Ques_Type = ("Perception (Q12,Q21)"),
  Chi_Square = fitMeasures(Q_per_cfa, "chisq"),
  df = fitMeasures(Q_per_cfa,"df"),
  P_value = fitMeasures(Q_per_cfa, "pvalue"),
  CFI = fitMeasures(Q_per_cfa, "cfi"),
  TLI = fitMeasures(Q_per_cfa, "tli"),
  RMSEA = fitMeasures(Q_per_cfa, "rmsea"),
  Raw_Alpha = Alpha2$total$raw_alpha,
  Std_Alpha = Alpha2$total$std.alpha,
  row.names = NULL
)


per_table


```


## Confirmatory Factor Analysis Results

```{r}

Summary_Table <- bind_rows(fam_table,per_table, awa_table)

Summary_Table 

```





## Subgroup Analysis


## SUMMARY BY GENDER AND CLASS


```{r Familiarity By Gender}
# by gender
# IMPORTANT: We restricted the working dataset to Female/Male at this point to create the student_data_MF subset for gender analysis, this exculdes students who declined to specify their gender or identified as non-banary.  

student_data_filtered <- student_data_filtered %>%
  filter(Q3 %in% c("Female", "Male"))

Q10_bygender <- student_data_filtered |> 
  group_by(Q3) |>
   filter(!is.na(class_group)) |>
  summarise(n= sum(!is.na(Q10_comp)),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE)) 

print(Q10_bygender)


# by class

Q10_class <- student_data_filtered |> 
  group_by(class_group) |>
  filter(!is.na(class_group)) |>
  summarise(n= sum(!is.na(Q10_comp)),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE))

print(Q10_class)


```




## BOXPLOT  FOR CLASS AND GENDER


```{r, warning=FALSE, message=FALSE}

student_data_filtered |>
  ggplot(aes(x = Q3, y = Q10_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = FALSE) +
  labs(
    title = "Q10 Composite Score by Gender",
    x = "Gender",
    y = "Q10 Composite Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)
  )



student_data_filtered |>
  ggplot(aes(x = class_group, y = Q10_comp )) +
  geom_boxplot(width = 0.6, alpha = 0.8) +
  labs(
    title = "Q10 Composite Score by Class Group",
    x = "Class Group",
    y = "Q10 Composite Score"
  )

student_data_filtered |> filter(!is.na(class_group), !is.na(Q10_comp))|>
 ggplot(aes(x = class_group, y = Q10_comp, fill = Q3)) +
  geom_boxplot(width = 0.6, alpha = 0.8, show.legend = TRUE) +
  labs(
    title = "Familiariy Composite Score by Class and Gender",
    x = "Class Group",
    y = "Familiariy Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "bottom"
  )


```

## SUMMARY


```{r Familirity (Q10) Summary}

Q10_Sum <- student_data_filtered |>
  summarise(n = sum(!is.na(Q10_comp)),
            Min = min(Q10_comp, na.rm = TRUE),
            Mean = mean(Q10_comp, na.rm = TRUE),
            SD = sd(Q10_comp, na.rm =TRUE),
            Max = max(Q10_comp, na.rm = TRUE))

print(Q10_Sum)




```
